### 2016-08-23 latex/stackoverflow.tex

% This is LLNCS.DOC the documentation file of
% the LaTeX2e class from Springer-Verlag
% for Lecture Notes in Computer Science, version 2.4
%\documentclass{llncs}
\documentclass{IEEEtran}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{listings}  
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{float}
\usepackage{array}
\usepackage{ragged2e}
\newcolumntype{P}[1]{>{\RaggedRight\hspace{0pt}}p{#1}}

\usepackage{etoolbox}
\makeatletter
\patchcmd{\@makecaption}
{\scshape}
{}
{}
{}
\makeatother

\lstdefinestyle{base}{
	language=C,
	emptylines=1,
	breaklines=true,
	basicstyle=\ttfamily\color{black},
	moredelim=**[is][\color{red}]{@}{@},
}

\def\tablename{Table}

\usepackage[justification=centering]{caption}

\newcommand{\squeezeup}{\vspace{-2.5mm}}
\newcommand\FIXME[1]{\textbf{FIXME: #1}}
%
\begin{document}

\onecolumn

\title{Code Overflow: \\
	An empirical study of online code reuse}
\maketitle

\section*{Raw statistics}
The results of running 2 clone detectors: Simian and NiCad, to detect clones between 144,075 Stackoverflow fragments (Java accepted answers) and 63 open-source projects in Qualitas dataset is presented below. There are 2 tools selected: Simian and NiCad. They are configured using three different settings: default settings, settings from FSE'13 paper, and settings from CloPlag experiment. Full Simian's parameter names can be found from the footnote\footnote{Simian's parameters: iChar = ignoreCharacters, iCurlB = ignoreCurlyBraces, iId = ignoreIdentifiers, iIdC = ignoreIdentifierCase, iMod = ignoreModifiers \newline iStrC = ignoreStringCase, iStr = ignoreStrings, iSbtNm = ignoreSubtypeNames, bSqBrck=balanceSquareBrackets}.

Manual investigation of Simian's clone report showed that there were problematic 11 fragments. These fragments generate false clone containing array initialisation. Hence, they were removed from the result set before analysis.

\begin{table*}[h]
	\centering
	\caption{Qualitas (2013-09-01r): Simian and NiCad clone results}
	\label{t_simian_raw_results}
	\resizebox{\columnwidth}{!}{%
	\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
		\hline
		\multirow{4}{*}{Statistics} 
		& \multicolumn{4}{c|}{Simian Default settings} 
		& \multicolumn{4}{c|}{Simian FSE'13 settings}
		& \multicolumn{4}{c|}{NiCad Default settings} 
		& \multicolumn{4}{c}{NiCad FSE'13 settings} \\ \cline{2-17} 
		& \multicolumn{4}{c|}{threshold=6} 
		& \multicolumn{4}{c|}{threshold=5}  
		& \multicolumn{4}{c|}{threshold=6, UPI=0.30} 
		& \multicolumn{4}{c}{threshold=5, UPI=0.2} \\ 
		& \multicolumn{4}{c|}{iCharC,iCurlB,iIdC,iMod,iStrC} 
		& \multicolumn{4}{c|}{iCurlB,iId,iIdC,iStr,iChar,iSbtNm,bSqBrck} 
		& \multicolumn{4}{c|}{rename=none,abstract=none} 
		& \multicolumn{4}{c}{rename=blind,abstract=literal} \\\cline{2-17}  
		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ 
		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ 
		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$
		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ \\ 
		\hline
		%Removed & 11 & -- & -- & -- & 0 & -- & -- & -- & -- & -- & -- & -- & -- & -- & -- & -- \\ %& 11 & -- & -- & -- \\
		Total  	& 654 	& 24,929 	& -- 	& --	& 1,175 & 16,957,362 & --	& --	& 985 	& 105,118 	& --	& --	& 12,084	& 113,557,298 & --		& -- \\
		Mean  	& --	& 38 		& 7.54 	& 0.27 	& --	& 14,444 	& 4.80 	& 0.28 	& --	& 107 		& 9.52 	& 0.25 	& -- 		& 9,397 	& 5.21 		& 0.20 \\
		Std Dev. & --  	& 87 		& 3.21 	& 0.22 	& --	& 281,747 	& 1.22 	& 0.18 	& --	& 198 		& 3.07 	& 0.18 	& --		& 12,098 	& 1.73 		& 0.16 \\
		Max  	& --	& 551 		& 49.00 & 0.94 	& --	& 9,599,676 & 18.00 & 0.89 	& --	& 1,792 	& 39.00 & 0.80 	& --		& 227,077 	& 44.00 	& 0.86 \\
		Min  	& --	& 1 		& 5.00 	& 0.01 	& --	& 1 		& 4.00 	& 0.02 	& --	& 1 		& 7.00 	& 0.02 	& --		& 1 		& 3.00 		& 0.01 \\
		Median  & --	& 3 		& 7.00 	& 0.23 	& --	& 22 		& 5.00 & 0.24 	& --	& 15 		& 8.00 	& 0.19 	& --		& 6,105 	& 5.00 		& 0.15 \\
		Mode  	& --	& 1 		& 7.00 	& 0.25 	& --	& 1 		& 4.00 	& 0.50 	& --	& 1 		& 8.00 	& 0.53 	& --		& 1 		& 4.00 		& 0.33 \\
		\hline
	\end{tabular} %
}
\end{table*}

%\begin{table*}[h]
%	\centering
%	\caption{Qualitas (2016-08-05): Simian and NiCad clone results (44 new + 19 old Qualitas projects)}
%	\label{t_simian_raw_results}
%	\resizebox{\columnwidth}{!}{%
%	\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
%		\hline
%		\multirow{4}{*}{Statistics} 
%		& \multicolumn{4}{c|}{Simian Default settings} 
%		& \multicolumn{4}{c|}{Simian FSE'13 settings} 
%		& \multicolumn{4}{c|}{NiCad Default settings} 
%		& \multicolumn{4}{c}{NiCad FSE'13 settings} \\ \cline{2-17} % & \multicolumn{4}{c}{CloPlag settings} \\
%		& \multicolumn{4}{c|}{threshold=6} 
%		& \multicolumn{4}{c|}{threshold=5} 
%		& \multicolumn{4}{c|}{threshold=6, UPI=0.30} 
%		& \multicolumn{4}{c}{threshold=5, UPI=0.2} \\  % & \multicolumn{4}{c}{threshold=5} \\
%		& \multicolumn{4}{c|}{iCharC,iCurlB,iIdC,iMod,iStrC} 
%		& \multicolumn{4}{c|}{iCurlB,iId,iIdC,iStr,iChar,iSbtNm,bSqBrck} 
%		& \multicolumn{4}{c|}{rename=none,abstract=none} 
%		& \multicolumn{4}{c}{rename=blind,abstract=literal} \\\cline{2-17}  %  & \multicolumn{4}{c}{iId} \\
%		% \hline
%		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ 
%		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$
%		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ 
%		& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ \\ % & Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ \\
%		\hline
%		%Removed & 11 	& -- 	& -- 	& -- 	& 0 & -- & -- & -- \\ %& 11 & -- & -- & -- \\
%		Total 	& 789 	& 61,164 & -- 	& -- 	&  170 	& 11,915,544 & -- 	& --  \\ % & 26,699 & 24,9702,821 & -- & --  \\
%		Mean 	& -- 	& 78 	& 7.32 	& 0.26 	& -- 	& 70,091 	& 4.76 	& 0.29 \\ % & -- & 9,356 & 4.74 & 0.26 \\
%		Std Dev. & -- 	& 208 	& 2.71 	& 0.21 	& -- 	& 703,788 	& 0.89 	& 0.19 \\ % & -- & 96,038	& 0.99 & 0.18 \\
%		Max 	& -- 	& 1,300 & 49 	& 0.94 	& -- 	& 9,065,774 & 9		& 0.83 \\ % & -- & 7,837,204 & 28 & 0.94 \\
%		Min 	& -- 	& 1 	& 5 	& 0.01 	& -- 	& 1 		& 4 	& 0.01 \\ % & -- & 1 & 4 & 0.002 \\
%		Median 	& -- 	& 2 	& 7 	& 0.20 	& -- 	& 17 		& 5 	& 0.26 \\ % & -- & 27 & 4 & 0.22 \\
%		Mode 	& -- 	& 1 	& 7 	& 0.33 	& -- 	& 1 		& 4 	& 0.33 \\ % & -- & 1 & 4 & 0.50 \\
%		\hline
%	\end{tabular} %
%}
%\end{table*}
%
%\begin{table*}[h]
%	\centering
%	\caption{Qualitas (2016-08-05): Simian and NiCad clone results (only 44 new Qualitas projects)}
%	\label{t_simian_raw_results}
%	\resizebox{\columnwidth}{!}{%
%		\begin{tabular}{l|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r|r}
%			\hline
%			\multirow{4}{*}{Statistics} 
%			& \multicolumn{4}{c|}{Simian Default settings} 
%			& \multicolumn{4}{c|}{Simian FSE'13 settings} 
%			& \multicolumn{4}{c|}{NiCad Default settings} 
%			& \multicolumn{4}{c}{NiCad FSE'13 settings} \\ \cline{2-17} % & \multicolumn{4}{c}{CloPlag settings} \\
%			& \multicolumn{4}{c|}{threshold=6} 
%			& \multicolumn{4}{c|}{threshold=5} 
%			& \multicolumn{4}{c|}{threshold=6, UPI=0.30} 
%			& \multicolumn{4}{c}{threshold=5, UPI=0.2} \\  % & \multicolumn{4}{c}{threshold=5} \\
%			& \multicolumn{4}{c|}{iCharC,iCurlB,iIdC,iMod,iStrC} 
%			& \multicolumn{4}{c|}{iCurlB,iId,iIdC,iStr,iChar,iSbtNm,bSqBrck} 
%			& \multicolumn{4}{c|}{rename=none,abstract=none} 
%			& \multicolumn{4}{c}{rename=blind,abstract=literal} \\\cline{2-17}  %  & \multicolumn{4}{c}{iId} \\
%			% \hline
%			& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ 
%			& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$
%			& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ 
%			& Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ \\ % & Frag.~ & $C_{\mathrm{pairs}}$ & $C_{\mathrm{SLOC}}$ & $C_{\mathrm{\%}}$ \\
%			\hline
%			%Removed & 11 	& -- 	& -- 	& -- 	& 0 & -- & -- & -- \\ %& 11 & -- & -- & -- \\
%			Total 	& 736 	& 60,678 & -- 	& -- 	&  166 	& 10,273,534 & -- 	& --  \\ % & 26,699 & 24,9702,821 & -- & --  \\
%			Mean 	& -- 	& 82 	& 7.26 	& 0.26 	& -- 	& 61,889 	& 4.77 	& 0.28 \\ % & -- & 9,356 & 4.74 & 0.26 \\
%			Std Dev. & -- 	& 214 	& 2.24 	& 0.21 	& -- 	& 607,074 	& 0.89 	& 0.18 \\ % & -- & 96,038	& 0.99 & 0.18 \\
%			Max 	& -- 	& 1,298 & 32 	& 0.94 	& -- 	& 7,720,970 & 9		& 0.83 \\ % & -- & 7,837,204 & 28 & 0.94 \\
%			Min 	& -- 	& 1 	& 5 	& 0.01 	& -- 	& 1 		& 4 	& 0.01 \\ % & -- & 1 & 4 & 0.002 \\
%			Median 	& -- 	& 2 	& 7 	& 0.20 	& -- 	& 16.5 		& 5 	& 0.26 \\ % & -- & 27 & 4 & 0.22 \\
%			Mode 	& -- 	& 1 	& 7 	& 0.33 	& -- 	& 1 		& 4 	& 0.33 \\ % & -- & 1 & 4 & 0.50 \\
%			\hline
%		\end{tabular} %
%	}
%\end{table*}

\begin{table}[h]
	\centering
	\caption{63 Qualitas projects (new versions checked on 2016-08-05)}
	\label{t:new_and_old}
	\resizebox{\columnwidth}{!}{%
	\begin{tabular}{l|r|r|c|c|p{5cm}|p{3cm}}
	\hline 
	Projects & Old version & New versions & Latest change & Repo. & License & Notes \\
	\hline
	ArgoUML & 0.34 & 0.35.1 & 15/01/2014 	& & Eclipse Public License v.1.0 \\
	Compiere & 330 & -- & 01/06/2010 		& git & GNU General Public License v.2.0 & No longer OSS \\
	FitLibrary & 20100806 & 20111026 & 12/07/2014 & & GNU General Public License v.2.0 \\
	JGroups & 2.10.0.GA & 3.6.10 & 27/06/2016 & & Apache License, v.2.0 \\
	Vuze & 4812 & 5721-20 & 12/05/2016 & & GNU General Public License v.2.0 &  \\
	antlr4 & 4.0 & 4.2.1 & 15/01/2014 & & BSD License \\
	aoisrc & 281 & 302 & 12/01/2015 & & GNU General Public License v.2.0 \\
	apache-ant & 1.8.4 & 1.10.0 & 09/04/2016 & git & Apache License, v.2.0 \\
	aspectj & 1.6.9 & 1.8.9 & 14/03/2015 & & Eclipse Public License 1.0 \\
	axion & 1.0-M2 & -- & 08/03/2013 & & Proprietary (BSD/Apache-style) \\
	batik & 1.7 & 1.8 & 16/03/2015 & & Apache License, v.2.0 \\
	c-jdbc & 2.0.2 & -- & 16/09/2005 & & GNU Lesser General Public License, v.2.1 \\
	castor & 1.3.1 & 1.4.2 & 20/06/2016 & git & Apache License, v.2.0 \\
	cayenne & 3.0.1 & 4.0.M4 &  27/07/2016 & git & Apache License, v.2.0 \\
	checkstyle & 5.1 & 7.2  & 04/08/2016 & git & GNU Lesser General Public License, v.2.1 \& Apache License, v.2.0 & \textit{Cli, Logging and Beanutils} packages are from the Apache Commons project. \\
	cobertura & 1.9.4.1 & 2.1.2 & 01/06/2015 & git & GNU General Public License v.2.0  \\
	colt & 1.2.0 & -- & 09/09/2014 & & Proprietary (CERN) \\
	columba & 1.4 & -- & 20/04/2007 & & Mozilla Public License 1.1  \\
	commons-collections & 3.2.1 & 4.1 & 25/11/2015 & & Apache License, v.2.0 \\
	db-derby & 10.6.1.0 & 10.12.1.1 & 13/09/2015 & & Apache License, v.2.0 \\
	displaytag & 1.2 & 2.0 & 17/08/2014 & svn & MIT License \\
	drawswf & 1.2.9 & -- & 02/04/2013 & & GNU General Public License v.2.0 \\
	drjava & 20100913-r5387 & 20140826-r5761 & 26/08/2014 & & BSD License \\
	eXoPortal & ??? & 4.4 & 24/06/2016 & git & GNU Lesser General Public License v.3.0 \& proprietary \\
	emma & 2.0.5312 & 2.0.5312 & 09/05/2013  & & Common Public License 1.0 \\
	findbugs & 1.3.9 & 3.0.1 & 06/03/2015 & & GNU Lesser General Public License, v.2.0 \\
	fit-java & 1.1 & -- & 04/06/2013 & & GNU General Public License v.2.0 \\
	freecol & 0.10.7 & 0.11.6 & 04/08/2016 & git & GNU General Public License v.2.0 &  \\
	freecs & 1.3.20100406 & -- & 22/04/2013 & & GNU General Public License v.3.0 &   \\
	freemind & 0.9.0 & 1.0.0 & 26/06/2016 & & GNU General Public License v.2.0+ &  \\
	galleon & 2.3.0 & 2.5.6 & 29/04/2013 & & GNU General Public License v.2.0 \\
	ganttproject & 2.0.9 & 2.8.1 & 05/08/2016 & & GNU General Public License v.3.0 &  \\
	geotools & 2.7-M3 & 16 & 02/08/2016 & git & GNU Lesser General Public License, v.2.0 \\
	hadoop & 1.0.0 & 2.6.4 & 11/02/2016 & & Apache License, v.2.0 \\
	heritrix & 1.14.4 & -- & 05/06/2013 & & GNU Lesser General Public License, v.2.1 \\
	hibernate & 4.2.2 & 5.2.2 & 04/08/2016 & git & GNU Lesser General Public License, v.2.1+ \\
	hsqldb & 2.0.0 & 2.3.4 & 16/05/2016 & svn & BSD License \\
	htmlunit & 2.8 & 2.24 & 27/07/2016 & svn & Apache License, v.2.0 \\
	iReport & 3.7.5 & -- & 28/05/2014 & & Affero GNU Public License, v.3.0 \\
	iText & 5.0.3 & 7.0.0 & 02/05/2016 & & Affero GNU Public License, v.3.0 \\
	informa & 0.7.0-alpha2 & -- & 07/11/2008 & & GNU Lesser General Public License, v.2.1 \& Apache Software License, v.1.1 \\
	ivatagroupware & 0.11.3 & -- & 27/02/2013 & & GNU General Public License v.2.0 \\
	jFin DateMath & R1\_0\_1 & -- & 25/04/2013 & & GNU General Public License v.2.0 \\
	jOggPlayer & 114s & -- &  15/04/2013 &  & GNU General Public License v.2.0 \\
	jag & 6.1 & 6.2 & 08/04/2013  & & GNU General Public License v.2.0 \& BSD License & BSD License is for libraries. \\
	james & 2.2.0 & 2.3.2.1 & 14/08/2015 & & Apache License, v.2.0 \\
	jasml & 0.10 & -- & 08/03/2013 & & Apache Software License \\
	jasperreports & 3.7.4 & 6.3.1 & 20/06/2016 & git & GNU Lesser General Public License, v.3.0 &  \\
	javacc & 5.0.0 & 7.0.0 & 07/07/2016 & svn & Proprietary (Sun) \\
	jboss (Wildfly) & 5.1.0.GA & 10.0.0 & 29/01/2016 & & GNU Lesser General Public License, v.2.1 & Renamed to Wildfly. \\
	jchempaint & 3.0.1 & 3.1.3 & 18/04/2016 & git & GNU Lesser General Public License, v.2.1+ &  \\
	jedit & 4.3.2 & 5.3.0 & 20/10/2015 & & GNU General Public License v.2.0 &  \\
	jena & 2.6.3 & 3.1.0 & 10/05/2016 & & Apache License, v.2.0 \\
	jext & 5.0 & -- & 18/08/2004 &  & GNU General Public License v.2.0 &  \\
	jfreechart & 1.0.13 & 1.2.0-ea1 & 08/05/2016 & svn & GNU Lesser General Public License, v.2.0 &   \\
	jgraph & 5.13.0.0 & 3.5.1.5 & 20/07/2016 & git & Proprietary (mxGraph License) \\
	jgraphpad & 5.10.0.2 & -- & 10/11/2006 & & GNU General Public License  \& GNU Lesser General Public License (derivatives) \\
	jgrapht & 0.8.1 & 0.9.2 & 03/04/2016 & & GNU Lesser General Public License, v.2.1 \& Eclipse Public License, v.1.0 \\
	jmoney & 0.4.4 & -- & 17/04/2016 & & GNU General Public License v.2.0 &  \\
	jparse & 0.96 & -- & 29/07/2004 & & GNU Lesser General Public License, v.2.1 \\
	jpf & 1.5.1 & ??? & 13/01/2012 & & Apache License, v.2.0 \\
	junit & 4.11 & 4.12 & 04/12/2014 & & Eclipse Public License, v.1.0 \\
	shiftone-jrat & 0.6 & 1-beta-1 & 25/04/2013 & svn & GNU Lesser General Public License, v.2.0 &  \\
	\hline 
\end{tabular} %
}
\end{table}

\newpage

\section*{Agreed Clone Pairs}
The investigation of the clone pairs using Bellon's \textit{good-match(p)} and \textit{ok-match(p)} criteria are listed in  Table \ref{t_agreed_good_clone_pairs} and Table \ref{t_agreed_ok_clone_pairs}. We use the threshold \textit{p} of 0.7 for both \textit{good-match} and \textit{ok-match}. Visualisation of \textit{good-match} common clone pairs between sets of parameter settings can be seen from Figure \ref{fig:venn4_pairs}. There are 14 \textit{good-match} pairs. The \textit{ok-match} subsumes the \textit{good-match} resulting in totally 528 common clone pairs.

However, for the 2 NiCad settings, we found some renaming and clustering errors which resulted in fewer projects having clone reports. For NiCad default settings, 6 projects had clustering failed errors: \textit{cayenne,checkstyle, db-derby, geotools, hibernate, iReport}. For NiCad FSE13 settings, 4 projects had renaming failed errors: \textit{Vuze, aspectj, eXoPortal, hibernate} and 13 projects had clustering failed errors: \textit{ArgoUML, castor, cayenne, checkstyle, db-derby, drjava, ganttproject, geotools, iReport, ivatagroupware, jasperreports, jboss, jchempaint} as depicted in Table \ref{tab:projects_missing}. So these projects are also missing from agreed clone pairs.

\begin{table}[H]
	\centering
	\caption{Qualitas projects successfully analysed by Simian and NiCad}
	\label{tab:projects_missing}
	\begin{tabular}{r|r|r|r}
		\hline
		Simian$_{df}$ & NiCad$_{df}$ & Simian$_{fse13}$ & NiCad$_{fse13}$ \\ \hline
		63	& 57 & 63 & 46 \\
		\hline
		& \textit{6 clustering failed} &  & \textit{13 clustering failed} \\
		\hline
		& cayenne   &	& cayenne \\
		& checkstyle &  & checkstyle \\
		& db-derby 	& & db-derby \\
		& geotools 	& & geotools \\
		& iReport 	& & iReport \\
		& hibernate & & ArgoUML \\
		&			& & castor \\
		&			& & drjava \\
		&			& & ganttproject \\
		&			& & ivatagroupware \\
		&			& & jasperreports \\
		&			& & jboss \\
		&			& & jchempaint \\
		\hline
		& \textit{renaming failed} & &  \textit{4 renaming failed} \\
		\hline
		&			& & Vuze \\
		&			& & aspectj \\
		&			& & eXoPortal \\
		&			& &  hibernate \\
		\hline
	\end{tabular}
\end{table}

\begin{table}[H]
	\centering
	\caption{1,415 agreed good-match(0.7) cloned pairs between Simian and Nicad (1,406 are unique)}
	\label{t_agreed_good_clone_pairs}
	\begin{tabular}{l|r|r}
		\hline
		& Simian$_{\mathrm{\textit{default}}}$ & Simian$_{\mathrm{\textit{fse13}}}$ \\ % & Simian$_{\mathrm{\textit{cloplag}}}$ \\
		\hline
		NiCad$_{\mathrm{\textit{default}}}$ & 9 & 6 \\ %      & 709             \\
		NiCad$_{\mathrm{\textit{fse13}}}$   & 12 & 1388 \\ % & 4680            \\
		%NiCad$_{\mathrm{\textit{cloplag}}}$ &             &               \\ % &                \\
		\hline
	\end{tabular}
\end{table}

%\begin{table}[H]
%	\centering
%	\caption{Duplicated agreed \textit{good}-pairs between pairs of Simian and NiCad settings}
%	\label{t_dup_agreed_good_clone_pairs}
%	\begin{tabular}{l|r|r|r|r}
%		\hline
%		& $S_dN_d$ & $S_dN_f$ & $S_fN_d$ & $S_fN_f$ \\ % & Simian$_{\mathrm{\textit{cloplag}}}$ \\
%		\hline
%		$S_dN_d$ & 9 &    &   &      \\%      & 709             \\
%		$S_dN_f$ & 9 & 12 &   &      \\ % & 4680            \\
%		$S_fN_d$ & 0 & 0  & 6 &      \\
%		$S_fN_f$ & 0 & 0  & 0 & 1388 \\	
%		%NiCad$_{\mathrm{\textit{cloplag}}}$ &             &               \\ % &                \\
%		\hline
%	\end{tabular}
%\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.5\linewidth]{venn4_pairs_good}
	\caption[Agreed clone pairs reported by Simian and NiCad with default and FSE'13 settings using good-match(0.7) criterion]{Agreed clone pairs reported by Simian and NiCad with default and FSE'13 settings using good-match(0.7) criterion}
	\label{fig:venn4_pairs}
\end{figure}

\begin{table}[H]
	\centering
	\caption{11,165 agreed \textit{ok}-match(0.7) cloned pairs between Simian and Nicad (10,879 are different, 286 are duplicates). They contain all \textit{good}-match pairs}
	\label{t_agreed_ok_clone_pairs}
	\begin{tabular}{l|r|r}
		\hline
		& Simian$_{\mathrm{\textit{default}}}$ & Simian$_{\mathrm{\textit{fse13}}}$ \\ % & Simian$_{\mathrm{\textit{cloplag}}}$ \\
		\hline
		NiCad$_{\mathrm{\textit{default}}}$ & 481 & 60 \\%      & 709             \\
		NiCad$_{\mathrm{\textit{fse13}}}$   & 418 & 10,206 \\ % & 4680            \\
		%NiCad$_{\mathrm{\textit{cloplag}}}$ &             &               \\ % &                \\
		\hline
	\end{tabular}
\end{table}

\begin{table}[H]
	\centering
	\caption{Duplicated agreed \textit{ok}-pairs between Simian and NiCad settings (10,879 are different, 286 are duplicates)}
	\label{t_dup_agreed_ok_clone_pairs}
	\begin{tabular}{l|r|r|r|r|r}
		\hline
		& $S_dN_d$ & $S_dN_f$ & $S_fN_d$ & $S_fN_f$ & Total \\ % & Simian$_{\mathrm{\textit{cloplag}}}$ \\
		\hline
		$S_dN_d$ & 214 	& 267 	& 0		& 0 		& 481 \\
		$S_dN_f$ & 267 	& 150 	& 0 	& 1 		& 418 \\ 
		$S_fN_d$ & 0 	& 0 	& 42 	& 18  		& 60 \\
		$S_fN_f$ & 0 	& 1 	& 18 	& 10,187 	& 10,206 \\	
		\hline
		Total 	 & 481  & 418	& 60	& 10,206	& 11,165 \\
		%NiCad$_{\mathrm{\textit{cloplag}}}$ &             &               \\ % &                \\
		\hline
	\end{tabular}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.5\linewidth]{venn4_pairs_ok}
	\caption[Agreed clone pairs reported by Simian and NiCad with default and FSE'13 settings using ok-match(0.7) criterion]{Agreed clone pairs reported by Simian and NiCad with default and FSE'13 settings using ok-match(0.7) criterion}
	\label{fig:venn4_pairs}
\end{figure}

\section*{Investigation of Missing A/B Clone Pairs Reported by Simian (df)}
We investigated the 41 clone pairs previously reported by Simian with default configurations and manually investigated. The 41 pairs were searched for in 4 new results sets: Simian$_{df}$, Simian$_{fse13}$, NiCad$_{df}$, NiCad$_{fse13}$. The investigation results are shown in Table \ref{tab:search}.

\begin{table}[H]
	\centering
	\caption{Results of matching the original 41 Simian(default) pairs in the pretty-printed result sets}
	\label{tab:search}
	\begin{tabular}{l|r|r}
		\hline 
		Settings & Found & Not found \\ 
		\hline 
		Simian$_{df}$  &  40 & 1* \\ 
		\hline 
		Simian$_{fse13}$ & 0 & 41  \\ 
		\hline 
		NiCad$_{df}$  & 17 & 24 \\ 
		\hline 
		NiCad$_{fse13}$ &  24 & 17 \\ 
		\hline 
	\end{tabular} 
\end{table}

The single missing Stackoverflow fragment (19051537\_0.java) (denoted by *) is one of the 11 false clones generated by Simian. It is removed from the results of the pretty-printed version because it is an outlier. The rest are missing because of different parameter settings.

\section*{Manual Investigation}

\subsection{Investigation of agreed \textit{good} and \textit{ok}-match clone pairs}
The classification scheme is described in Table \ref{tab:classification_scheme} and the classification results are shown in Table \ref{tab:classification}. We have manually investigated all of the 1,406 \textit{good-match} ones reported by agreement of four different Simian and NiCad settings.  However, for the \textit{ok-match}, we could not investigate all of the 10,879 pairs manually.  According to the distribution of category from \textit{good-match} results, we can see that Simian$_{\textrm{\textit{fse13}}}$--NiCad$_{\textrm{\textit{fse13}}}$ produces a large number, 1,387, of false positive results (D, E, and F). Thus, we decided to leave them out of the manual investigation of \textit{ok-match} pairs. There are totally 664 \textit{ok-match} pairs that were investigated. The 57 true positive pairs found are combinations of 10 unique Stackoverflow fragments, and 13 unique Qualitas Java files.

Since we are not certain about the direction of copying in the B-classified pairs, we checked the modification time of each Java file in Qualitas project and compare it to the timestamp of Stackoverflow answers. We found that all Stackoverflow code fragments were posted after their respectively similar Java files in Qualitas project. This means that the copying can only be either (1) Q $\rightarrow$ S or (2) from a third source to both S and Q independently.

\subsection{Investigation of the unique clone pairs}
In the preliminary stage of our experiment, we found that there are 41 Stackoverflow fragments reported by Simian with default configurations. However, only 10 of them appear in the new results using tool's agreement. Thus, we further investigated the clone pairs reported by Simian and NiCad but \textit{without} an agreement. With our 4 settings, we investigated the largest 172 clone pairs according to their average SLOC (larger than or equals to 15 SLOC). These clone pairs are pairs of Stackoverflow fragments and Java files from Qualitas projects that NEVER been looked at before in the previous investigation. The reason for having this criteria is that we found a lot of duplicate clone pairs with the same classifications from either Stackoverflow fragments or Qualitas files. The classification results are shown in Table \ref{tab:classification}, \ref{tab:good_classification} and Table \ref{tab:good_classification}.

\begin{table}[H]
	\centering
	\caption{Classification results of 1,406 \textit{good-} and 664 \textit{ok}-matches (excluding the subsumed \textit{good}-match pairs).}
	\label{tab:classification}
	\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
		\hline 
		Classification & A & A' & B & C & D & E & F & Total \\ 
		\hline 
		\multirow{1}{*}{\textit{good-match(0.7)}} & 1 & 0 & 1 & 3 & 26 & 6 & 1,369 & 1,406 \\ \cline{2-8}
		%& 7.14\%	& & 14.29\%	& 14.29\%	& 14.29\%	& 50.00\%	& 0.00\%	& 100.00\% \\ 
		\hline
		\multirow{1}{*}{\textit{ok-match(0.7)}} & 8  & 0 & 36 & 8 & 474 & 57 & 81 & 664 \\ \cline{2-8}
		%& 0.78\%	& & 8.24\%	&	1.57\%	&	68.82\%	&	5.49\%	&	15.10\%	&	100.00\%  \\ 
		\hline
		\multirow{1}{*}{??? indv. pairs} &  & &  &  &  &  &  &  \\
		\hline
		\multirow{3}{*}{TP/FP} & \multicolumn{4}{c|}{57} & \multicolumn{3}{c|}{2,013} & 2,070 \\ 
	    & \multicolumn{4}{l|}{\textit{10 unique SO fragments}} & \multicolumn{3}{c|}{} & \\
	    & \multicolumn{4}{l|}{\textit{13 unique Q files}} & \multicolumn{3}{c|}{} & \\
		\hline
	\end{tabular} 
\end{table}

\begin{table}[H]
	\centering
	\caption{Distribution of classification category A--F according to \textit{good}-match pairs}
	\label{tab:good_classification}
\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
	\hline 
	Category   & A   & 	A'   & 	B   & C   & D   &	E   &	F   & Total  \\
	\hline
	Simian$_{df}$--NiCad$_{df}$   & 1 & 0 & 1 & 0 & 0 & 4 & 0 & 6 \\
	Simian$_{df}$--NiCad$_{\textrm{\textit{fse13}}}$   & 0 & 0 & 0 & 3 & 1 & 1 & 1 & 6 \\
	Simian$_{\textrm{\textit{fse13}}}$--NiCad$_{df}$   & 0 & 0 & 0 & 0 & 7 & 0 & 0 & 7 \\
	Simian$_{\textrm{\textit{fse13}}}$--NiCad$_{\textrm{\textit{fse13}}}$   & 0 & 0 & 0 & 0 & 18 & 1 & 1,368 & 1,387 \\
	\hline
	Total   &   1   &   0   &   1   &   3   &   26   &   6   & 1,369  & 1,406 \\
	\hline
\end{tabular} 
\end{table}

\begin{table}[H]
	\centering
	\caption{Distribution of classification category A--F  according to the \textit{ok}-match pairs}
	\label{tab:ok_classification}
	\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
		\hline 
		Category   																										& A   	& 	A' 	& 	B  & C	   & D   	&	E   &	F   & Total  \\
		\hline
		Simian$_{df}$--NiCad$_{df}$   																	& 3 	& 0 	& 17	& 1 	& 428  & 6 		& 7 	&  462 \\
		Simian$_{df}$--NiCad$_{\textrm{\textit{fse13}}}$   									& 5 	& 0 	& 18 	& 7 	& 17 	 & 51 	  & 50 		&  148 \\
		Simian$_{\textrm{\textit{fse13}}}$--NiCad$_{df}$   									& 0 	& 0 	& 1 	& 0 	 & 29 	  & 0 		& 24 	& 54 \\
		%Simian$_{\textrm{\textit{fse13}}}$--NiCad$_{\textrm{\textit{fse13}}}$   & 0 	& 0 	& 3 	& 0 	& 6 	  & 0 		& 0 	& \\
		\hline
		Total   &   8  &   0   &  36   &  8   &   474   &   57   & 81  & 664 \\
		\hline
	\end{tabular} 
\end{table}

%\begin{table}[H]
%	\centering
%	\caption{Breakdowns of the biggest 172 individual clone pairs}
%	\label{tab:classification_breakdown}
%	\begin{tabular}{|l|r|r|r|r|r|r|r|}
%		\hline 
%		Classification & A & B & C & D & E & F & Total \\ 
%		\hline 
%		Simian$_{df}$ & 16 & 7 & 3 & 7 & 4 & 3 & 40 \\ 
%		\hline
%		Simian$_{fse13}$ & 1 & 0 & 0 & 1 & 0 & 53 & 55 \\
%		\hline
%		NiCad$_{df}$ & 0 & 0 & 0 & 13 & 0 & 22 & 35 \\
%		\hline
%		NiCad$_{fse13}$ & 0 & 0 & 0 & 0 & 1 & 41 & 42 \\
%		\hline
%		Total & 17 & 7 & 3 & 21 & 5 & 119 & 172 \\
%		\hline
%	\end{tabular} 
%\end{table}

\begin{table}[H]
	\centering
	\caption{Classifications of clone creation}
	\label{tab:classification_scheme}
	\begin{tabular}{|c|p{13cm}|}
		\hline 
		Category & Descriptions \\ 
		\hline 
		A & Code in Stackoverflow is copied from Qualitas (Q $\rightarrow$ S). \\ 
		\hline 
		A' & Code in Qualitas is copied from Stackoverflow (S $\rightarrow$ Q). \\ 
		\hline 
		B & Code is copied either from each other or a third source (unknown) (S $\leftrightarrow$ Q $\vee$ (T $\rightarrow$ S $\wedge$ T $\rightarrow$ Q)).
		\\ 
		\hline 
		C & Code in both places are copied from a third source T (known) (T $\rightarrow$ S $\wedge$ T $\rightarrow$ Q).
		\\ 
		\hline 
		D & Code is a boiler-plate or IDE auto-generated.
		\\ 
		\hline 
		E & Code in both places initialise a similar/the same object; extend the same class/its subclass; implement the same interface.
		\\ 
		\hline 
		F & Accidental similarity, false clone \\ 
		\hline 
	\end{tabular} 
\end{table}

\newpage

\section*{Simian's Parameters}

We have carefully investigated the effects of the Simian's parameter \texttt{-balanceSquareBrackets+}. I found that it works in the expected way of handling a pair of brackets (\texttt{[},\texttt{]}) that span over multiple lines. For example, the two code fragments in Figure \ref{fig:two_frags} would match by having \texttt{-balanceSquareBrackets+} enabled.

\noindent\begin{figure}[H]
	\scriptsize
	\begin{lstlisting}[frame=single,style=base]
1           public class MagicSquare {                             public class MagicSquare2 {
2               private int[][] square;                                @private int[@
3               private boolean[] possible;                                        @][@
4               private int totalSqs;                                              @] square;@
5               private int sum;                                       private boolean[] possible;
6               private int numsquares;                                private int totalSqs;
7               public static void main ( String[] args ) {            private int sum;
8               MagicSquare m = new MagicSquare ( 3 );                 private int numsquares;
9                                                                      public static void main ( String[] args ) {
10                                                                     MagicSquare m = new MagicSquare ( 3 );
	\end{lstlisting}
	\caption{Two identical fragments with only differences in locations of the square brackets. All 7 lines are reported by Simian if \small\texttt{-balanceSquareBrackets+} \normalsize is enabled. If not, the clone pairs is reported as \newline (MagicSquare.java [3,8], MagicSquare2.java [5,10]).} 
	\label{fig:two_frags}
\end{figure}

However, the -balanceSquareBrackets+ parameter only works on a small testing environment having toy programs or only small pairs from the full datasets. It does not work with the full complete set of 144,075 Stackoverflow fragments and Qualitas projects. Please find the summary of all the testing scenarios in Table \ref{tab:summary}. %I found that this pair is missing if I turned on \texttt{-balanceSquareBrackets+}.

\begin{table}[H]
	\caption{Simian's \texttt{-balanceSquareBrackets+} (\texttt{-bsb+}) is observed to have unpredictable behaviours when running against big datasets. $\textrm{\textit{CP}}_2$ means the reported clone(s) do not contain lines having dislocated brackets ($L_b$) (i.e. $\textrm{\textit{CP}}_2 = \textrm{\textit{CP}}_1 - L_b$).}
	\label{tab:summary}
	\centering
	\small\begin{tabular}{l|p{6cm}|c|c|c}
		\hline 
		\multirow{2}{*}{\textbf{Project 1}} & \multirow{2}{*}{\textbf{Project 2}} & \textbf{Dislocated} & \multirow{2}{*}{\textbf{-bsb+}} & \textbf{Clones pair} \\ 
		& & \textbf{brackets?} & & \textbf{reported} \\
		\hline
		\hline
		\multicolumn{5}{l}{\textit{Only run Simian against the pair}} \\
		\hline
		MagicSquare.java & MagicSquare\_exact\_copy.java & no & 0,1 & $\textrm{\textit{CP}}_1$ \\
		MagicSquare.java & MagicSquare2.java & yes & 0 & $\textrm{\textit{CP}}_2$ \\ 
		MagicSquare.java & MagicSquare2.java & yes & 1 & $\textrm{\textit{CP}}_1$ \\
		\hline
		%\hline
		%\multicolumn{5}{c}{\textit{Only run Simian against the pair}} \\
		%\hline
		stackoverflow/4298836\_0.java & Qualitas/aoisrc281/../ExprModule.java & no & 0 & $\textrm{\textit{CP}}_3$ \\ 
		stackoverflow/4298836\_0.java & Qualitas/aoisrc281/../ExprModule.java & no & 1 & $\textrm{\textit{CP}}_3$ \\ 
		stackoverflow/4533682\_1.java & Qualitas/cobertura-1/../TouchCollector.java & no & 0 & $\textrm{\textit{CP}}_4$ \\ 
		stackoverflow/4533682\_1.java & Qualitas/cobertura-1/../TouchCollector.java & no & 1 & $\textrm{\textit{CP}}_4$ \\ 
		\hline 
		\hline
		\multicolumn{5}{l}{\textit{Run Simian against the complete stackoverflow data and the project}} \\
		\hline
		stackoverflow/4298836\_0.java & Qualitas/aoisrc281/../ExprModule.java & no & 0 & $\textrm{\textit{CP}}_3$ \\ 
		\cellcolor{red!10}stackoverflow/4298836\_0.java & \cellcolor{red!10}Qualitas/aoisrc281/../ExprModule.java & \cellcolor{red!10}no & \cellcolor{red!10}1 & \cellcolor{red!10}-- \\ 
		\hline
		stackoverflow/4533682\_1.java & Qualitas/cobertura-1/../TouchCollector.java & no & 0 & $\textrm{\textit{CP}}_4$ \\ 
		\cellcolor{red!10}stackoverflow/4533682\_1.java & \cellcolor{red!10}Qualitas/cobertura-1/../TouchCollector.java & \cellcolor{red!10}no & \cellcolor{red!10}1 & \cellcolor{red!10}-- \\
		\hline 
	\end{tabular} 
\end{table}

\end{document}