### 2017-02-18 cloneoverflow-latex/cloneoverflow.tex
(Section 2) merged the Experimental Framework and Experimental Setup together.

\documentclass[sigconf,review, anonymous]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{listings}  
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\usepackage{multirow}
%\usepackage[table]{xcolor}
\usepackage{float}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{cleveref}
\usepackage{subcaption}
\usepackage{tikz}

\newcolumntype{P}[1]{>{\RaggedRight\hspace{0pt}}p{#1}}

\lstset{language=Java,tabsize=2,basicstyle=\ttfamily\scriptsize}

\newcommand{\squeezeup}{\vspace{-0.5mm}}

\usepackage{etoolbox}
\makeatletter
\patchcmd{\@makecaption}
{\scshape}
{}
{}
{}
\makeatother

\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}

\lstdefinestyle{base}{
	language=C,
	emptylines=1,
	breaklines=true,
	basicstyle=\ttfamily\color{black},
	moredelim=**[is][\color{red}]{@}{@},
}

\newcommand\FIXME[1]{\textbf{FIXME: #1}}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[ESEC/FSE 2017]{11th Joint Meeting of the European Software Engineering Conference and the ACM SIGSOFT Symposium on the Foundations of Software Engineering}{4--8 September, 2017}{Paderborn, Germany}
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\begin{document}
	\title{Are We Reusing Outdated and Licensed Code from Stack Overflow?}
	%\titlenote{Produces the permission block, and
	%  copyright information}
	%\subtitle{Extended Abstract}
	%\subtitlenote{The full version of the author's guide is available as
	%  \texttt{acmart.pdf} document}

\author{Chaiyong Ragkhitwetsagul,~Matheus Paixao,~Jens Krinke}
\affiliation{%
	\institution{CREST, University College London}
	%  \streetaddress{P.O. Box 1212}
	\city{London, UK} 
	%  \state{Ohio} 
	%  \postcode{43017-6221}
}
%\email{{ucabagk,matheus.paixao.14,j.krinke}@ucl.ac.uk}

\author{Giuseppe Bianco,~Rocco Oliveto}
\affiliation{%
	\institution{Università degli Studi del Molise}
	%  \streetaddress{P.O. Box 1212}
	\city{Campobasso, Italy} 
	%  \state{Ohio} 
	%  \postcode{43017-6221}
}

\begin{abstract}
	We studied online code clones on the popular programming Q\&A website, Stack Overflow, and their issues on outdated code and software licensing violation. Online code clones are code fragments that are copied from software projects or online sources to Stack Overflow as examples. They can be discovered by using code clone detectors, yet they are inherently different from normal code clones in software projects in several aspects. Due to an absence of checking mechanism after the code has been copied to Stack Overflow, the online code clones may suffer from being outdated and violating the original software license. %Moreover, we also found that majority of online code clones do not contain the license of their originals. 
	We designed the experiment incorporating several state-of-the-art tools and techniques to automatically extract, prioritise, and filter the clones. As a result, we discovered 51,926 non-trivial online clone candidates between 144,064 Java code snippets on Stack Overflow and 111 Java open source projects in the curated Qualitas corpus. The manual study of the 3,632 online clone pairs revealed that 64 clones on Stack Overflow are outdated. Furthermore, we found 357 Stack Overflow code snippets that do not contain software license of their originals or have conflicting licenses. These snippets are questionable for reuse. Our study is among the first to address the problem of outdated and license-violating code snippets on Stack Overflow.
\end{abstract}

\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10011007.10011006.10011072</concept_id>
	<concept_desc>Software and its engineering~Software libraries and repositories</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10011074.10011111.10011113</concept_id>
	<concept_desc>Software and its engineering~Software evolution</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Software libraries and repositories}
\ccsdesc[300]{Software and its engineering~Software evolution}

% We no longer use \terms command
%\terms{Theory}

\keywords{clone detection, Stack Overflow, outdated code, software licensing}

\maketitle

\section{Introduction}

%The popularity of the Internet encourages tremendous amount of source code being shared online. 
Stack Overflow is a popular online programming community with 6.3
million users. It allows programmers to ask questions and give answers
to programming problems. The website has found to be useful for
software
development~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014,
  Stolee2014,Subramanian2013,Diamantopoulos2015,Treude2016} and also
valuable for educational purposes~\cite{Nasehi2012}. On Stack
Overflow, each conversation contains a question and a list of
answer(s). The answers frequently contain at least one code snippet as
a solution to the question asked. We found that the code snippets are
usually not authored directly on the Stack Overflow website but copied
from another location. An answer's snippet could be copied and modified
from a code snippet in the question, copied from the answerer's own
code or from other locations including open source software (OSS)
systems. The process of posting and answering questions on Stack
Overflow, which involves reusing of source code, can be considered as
code cloning.

Code cloning is an activity of reusing source code by copying and
pasting. It normally occurs in software development and account from
7\% to 23\% of source code in typical software
systems~\cite{Bellon2007}. The benefits and drawbacks of clones are
still controversial. Several authors state that clones lead to bug
propagations and software maintenance issues~\cite{Kamiya2002}, while
some others suggest that clones are not harmful and can even be
beneficial~\cite{Saini2016,Kapser2006}.

Code cloning can also have side effects such as violating software
licenses or introducing software vulnerabilities. Carelessly cloning
code from one project to another project with a different license may
cause a software licensing violation~\cite{German2009}. This also
happens within the context of online Q\&A websites such as Stack
Overflow. An et al.~\cite{An2017} showed that 1,279 cloned snippets
between Android apps and Stack Overflow potentially violate software
licenses. Security is also among the main concerns when code is copied
from an online source. Stack Overflow helps developers to solve
Android programming problems more quickly than other resources while,
at the same time, offers less secure code than books or the official
Android documentation~\cite{Acar2016}.

%\FIXME{Find the right location for this-->} Additionally, outdated and vulnerable third-party libraries from well-known open source projects are cloned into in a large number of open source systems~\cite{Xia2014}.

We call code snippets that are copied from software systems to online
Q\&A websites (such as Stack Overflow), and vice versa as ``online
code clones'' (sometimes only ``clones'' for brevity). There are two
directions in creating online code clones: (1) code is cloned from a software
project to a Q\&A website as an example; or (2) code is cloned from a
Q\&A website to a software project to obtain a functionality, perform
a particular task, or fixing a bug.
% and (3) code is transferred from one software project to another via a Q\&A website; 
Similar to classic clones, online code clones can lead to license
violations, bug propagation, introduction of vulnerabilities, and
re-use of outdated code. Unfortunately, online clones are more
difficult to locate and fix since the search space in online code
corpora is larger and no longer confined in a local repository.

In this study, we are interested in locating online code clones on
Stack Overflow that are cloned from open source projects and study
them on two aspects: \textbf{outdated code and license conflicts.}

\begin{figure*}
	\begin{lstlisting}
          /* Code in Stack Overflow post ID 22315734 */                  /* WritableComparator.java (2016-09-26) */
          public int compare (byte[] b1,int s1,int l1,                   public int compare(byte[] b1,int s1,int l1,
                              byte[] b2,int s2,int l2) {                                    byte[] b2,int s2,int l2) {
            try {                                                          try {
              buffer.reset(b1,s1,l1); /* parse key1 */                       buffer.reset(b1,s1,l1); /* parse key1 */
              key1.readFields(buffer);                                       key1.readFields(buffer);
              buffer.reset(b2,s2,l2); /* parse key2 */                       buffer.reset(b2,s2,l2); /* parse key2 */
              key2.readFields(buffer);                                       key2.readFields(buffer);
            } catch (IOException e) {                                        buffer.reset(null,0,0); /* clean up reference */
              throw new RuntimeException(e);                               } catch (IOException e) {
            }                                                                throw new RuntimeException(e);
            return compare(key1,key2); /* compare them */                  }
          }                                                                return compare(key1, key2); /* compare them */
	                                                                       }
	\end{lstlisting}
	\caption{A motivating example of the two code fragments of
          {\small\texttt{WritableComparator.java}}. The one from the
          Stack Overflow post 22315734 (left) is outdated when compared to
          its latest version in the \textsf{hadoop} code base
          (right).}
	\label{fig:before-after}
\end{figure*}

A motivating example of outdated online code clones can be found in an
answer to a Stack Overflow question regarding how to implement
{\small{\texttt{RawComparator}}} in
\textsf{hadoop}\footnote{\url{http://stackoverflow.com/questions/22262310}}.
\Cref{fig:before-after} shows a code snippet embedded as a part of the
accepted answer on the left. The snippet shows how \textsf{hadoop}
implements the {\small{\texttt{compare}}} method in its
{\small{\texttt{WritableComparator}}} class. The code snippet on the
right shows another version of the same method, but at this time
extracted from the latest version (as of 26 September 2016) of
\textsf{hadoop}. We can see that they are highly similar except a line
containing {\small{\verb|buffer.reset(null,0,0);|}} which was added in
the latest version. The added line is intended for cleaning up the
reference in the {\small{\verb|buffer|}} variable. While this change
has already been introduced into the {\small{\texttt{compare}}} method
in the latest version of \textsf{hadoop}, the code example in Stack
Overflow post is still unchanged. This example shows that there can be
inconsistencies between online code clones and their originals leading
users of Stack Overflow to reuse outdated code.
%The code snippet on Stack Overflow can be outdated. %Since reusing source code from Stack Overflow is considered a common practice nowadays, the scale of online code cloning is more than intra or inter project clone in a local code bases.
%This is an emerging and challenging problem. Since studies in this area are still limited, we aim to gain more insight of the problem in this study. %We are interested to gain more insights of online code clone in this study.

While research has mostly focused on reusing code snippets \emph{from}
Stack Overflow (e.g.~\cite{Keivanloo2014,An2017,Yang2016}), less work
has been conducted on finding the origins of code examples copied
\emph{to} Stack Overflow and problems of reusing them such as outdated
code and software licensing violations. This paper makes the following
primary contributions:

\vspace{0.5ex}%
\noindent\textbf{1.~A manual study of online code clones:} 
We used two clone detection tools to discover 315,786,118 similar code
snippet pairs, followed by a manual classification of 3,632 candidate
clone pairs between Java code snippets obtained from Stack Overflow's
accepted answers and 111 Java open source projects from the curated
Qualitas corpus~\cite{QualitasCorpus}.

\vspace{0.5ex}%
\noindent\textbf{2.~An investigation of outdated and license-violating
  online clones on Stack Overflow:} Our study shows that from the
3,632 online clones, at least 668 likely have been copied from open
source projects or external online sources to Stack Overflow,
potentially violating software licenses. For 112 of them, we found
evidence that they have been copied from a specific Qualitas project
and that 64 of them are meanwhile outdated.
% and questionable for being reused.

\vspace{0.5ex}%
\noindent\textbf{3.~An online code clone oracle:} The 3,632 manually
investigated and validated online clone pairs are available for
download\footnote{https://to.be.disclosed.after.review} and can be
used as a clone oracle.

\section{Empirical Study}

We performed an empirical study of online code clones between Stack
Overflow and 111 Java open source projects to answer the following
research questions: \\
%
\textbf{RQ1 (online code clones):} \textit{To what extent is source
  code cloned between Stack Overflow and open source projects?} We
quantitatively measured the number of online code clones between Stack
Overflow and open source projects to understand the scale of the
problem. \newline 
%
\textbf{RQ2 (Patterns of online code clones):} \textit{Why do online
  code clones occur?} We categorised online clones into seven
categories allowing insights to why online code clones are created.
\newline
%
\textbf{RQ3 (Outdated online code clones):} \textit{Are
  online code clones up-to-date compared to their counterparts in the
  original projects?} We were interested in the outdated Stack
Overflow code examples since users are potentially reusing
them. \newline
%
\textbf{RQ4 (Software licensing violation):} \textit{Do
  licensing conflicts occur between Stack Overflow clones and their
  originals?} We investigated whether the reuse of online code clones
can cause software developers to violate licenses.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{exp_framework_new}
  \caption{Experimental framework}
  \label{fig:exp_framework}
\end{figure}

%\subsection{Experimental Setup}
\vspace{1ex}
To answer these four research questions, we perform an experiment to
locate and study the online code clones between Stack Overflow and open source
projects. The experimental framework is depicted in \Cref{fig:exp_framework}. 
We discuss the experimental setup below.

%\subsection{Experimental Setup}

%We followed the experimental framework by processing the source code from the two data sets, configuring and running the clone detectors, and analysing the clones. We discuss each of them in details below.

\subsection{Datasets}

We reply on two data sets in this study: Java code snippets on Stack Overflow and open source projects from the Qualitas corpus.

\subsubsection{Stack Overflow} we extracted Java code snippets from a snapshot of a Stack Overflow dump\footnote{https://archive.org/details/stackexchange} in January 2016. The archived dump has a size of 9 gigabytes. The data dump was in XML format containing information of \textbf{Posts} (questions and answers) and supporting data such as user accounts and timestamps of the posts. We were interested in code snippets embedded in posts which were located between {\small\texttt{<code></code>}} tags. A Stack Overflow thread contains a post of a question and several posts of answers where every answer has voting score from its users. An answer can also be marked as \textbf{accepted answer} by the questioner if the solution fixes his/her problem. We collected Java code snippets using two criteria. First, we were only interested in code snippets with at least six lines. Snippets that are smaller than six lines are usually spurious clones~\cite{Bellon2007}. Second, we only focused on code snippets in accepted answers. We chose the snippets in accepted answers because they actually solved the problems in the questions. Moreover, they are always displayed just below the questions which makes them more attractive to be reused than other answers. Each snippet was extracted from the dump and saved to a file. 
%using its post ID as the file's name. If an accepted answer had more than one code snippet, we appended an indexing number, starting from zero, after the post ID (e.g. 45051109\_0, and 45051109\_1.). Lastly, we added {\small\texttt{.java}} extension to the file's names so that clone detectors could recognise them. 
We obtained 144,064 Java code snippets which contain 2,347,269 lines of Java source code excluding comments and blank lines\footnote{Measured by cloc: https://github.com/AlDanial/cloc}.

\textbf{Open source systems}: we selected the established corpus for an empirical software engineering study called \textbf{Qualitas}~\cite{QualitasCorpus} for this study. It is a curated Java corpus that has been used in several software engineering studies~\cite{Taube-Schock2011,Beckman2011,Vasilescu2011,Omar2012}. The projects in the corpus represent various domains of software systems ranging from programming languages to visualisation~\cite{QualitasCorpus}. We selected the 20130901r release of Qualitas corpus containing 112 Java open source projects. This release contains projects with releases no later than 1st September 2013. We chose a snapshot late back in 2013 since we are interested in online code clones in the direction from open source projects to Stack Overflow. The 20130901r snapshot provides Java code that is at least 3 years old from the time of the experiment (January to December 2016). The time difference is sufficiently long for a number of code snippets to be copied onto Stack Overflow. Out of 112 Qualitas projects, there is one project, \textsf{jre}, that does not contain Java source code due to its licensing limitation~\cite{QualitasCorpus} so it is removed from the study. This resulted in a total of 111 projects analysed in the study. As shown in \Cref{tab:datasets}, the 111 Qualitas project have 160,937 Java files containing 19,086,883 lines of code. %Details of the 109 Qualitas projects and their licenses are listed in Table \ref{t:new_and_old}.

\begin{table}
	\centering
	\caption{Stack Overflow and Qualitas datasets}
	\label{tab:datasets}
	\small
	\begin{tabular}{l|r|r}
		\hline 
		Dataset & No. of files & SLOC \\
		\hline
		Stack Overflow & 144,064 & 2,346,267 \\ 
		\hline 
		Qualitas &  160,937 & 19,086,883 \\ 
		\hline 
	\end{tabular} 
\end{table}

\subsection{Clone Detectors}
There is a number of restrictions in terms of choosing clone detection tools for this study. The main restriction is due to nature of code snippets posted on Stack Overflow, most of them are not complete Java classes or methods. Hence, the tool must be flexible enough to process code snippets that are neither a complete block nor compilable. Moreover, since the amount of code that has to be processed is in a scale of millions line of code (as shown in \Cref{tab:datasets}), a clone detector must be scalable enough to successfully complete the execution and report clones in a reasonable amount of time. We have tried 5 state-of-the-art clone detectors including Simian~\cite{simian}, NiCad~\cite{Cordy,Roy2008}, CCFinder~\cite{Kamiya2002}, iClones~\cite{Gode2009}, and DECKARD~\cite{Jiang2007a} against Stack Overflow and Qualitas datasets. CCFinder, iClones, and DECKARD failed with execution errors after running for couple of hours. Thus, we removed them from the study. Simian and NiCad completed the detection with success. We found that both of them are also flexible enough to handle code with incomplete methods or classes. %So, we decided to use both of them.

\textbf{Simian} is a text-based clone detector that locates clones at line-level granularity and has been used extensively in several clone studies~\cite{Ragkhitwetsagul2016, Wang2013, Mondal2011, Cheung2015, Krinke2010}. It is a command-line tool which enables us to automate the detection. Furthermore, it offers normalisation of variable names and literals (strings, and numbers) which enable Simian to detect literal clones (type-1) and parameterised clones (type-2). \textbf{NiCad} is also a text-based clone detector which detects clones at either method- or block-level granularity. It can detect clones of type-1, 2 up to type-3 (clones with added/removed/relocated/changed statements) and is also used in several empirical clone studies~\cite{Roy2008, Ragkhitwetsagul2016, Svajlenko2014, Wang2013, Mondal2011, Sajnani2016}. It utilises TXL~\cite{Cordy2006} for parsing and pretty-printing source code. It also provides code normalisation by variable renaming and abstraction. We use a variant of NiCad called \textit{nicadcross}. It offers the same functionalities as the original NiCad but is specialised for detecting code clones between two systems. NiCad is also a command-line tool which makes it suitable for automation.

%We processed two datasets, 144,064 Java code snippets from accepted answers on Stack
%Overflow and 111 open source projects from the Qualitas corpus. 
%The Java code snippets were extracted from Stack Overflow posts. 
We prepared the Java code in both datasets by removing comments and
pretty-printing to increase the accuracy of clone detection. Then, we
deployed two clone detection tools to locate clones between the two datasets.
Due to the scalability of Simian and NiCad, we partitioned the input
and run the tools multiple times. Each run was composed of the whole
Stack Overflow data and a single Qualitas project. We repeated the
process until we covered all 111 projects.
We then converted the clone reports to the General Clone Format
(GCF)~\cite{Wang2013} and combined the 111 reports into a single GCF
file.
Simian did not provide an option to detect cross-project clones
between two locations. Hence the Simian GCF clone report was filtered
to contain only clone pairs between Stack Overflow and Qualitas
projects, removing all clone pairs within either Stack Overflow or
within Qualitas. NiCad provides an option to detect clones from two
locations so no filtering was needed.

\subsection{Prioritise Clone Candidates Using Agreement}
The number of clones detected in large-scale datasets can be huge. In our study, there are in total 315,786,118 clone pairs reported. It is infeasible for humans to manually validate them all. One can do a random sampling of clones. However, due to a very high number of false positives, they may end up looking mostly at the false clones. Therefore, we adopted the idea of \textbf{clone agreement} which has been used in clone research studies~\cite{Wang2013,Funaro2010,cr2016ssbse} in situations where a clone oracle is missing or impossible to establish. Clone pairs agreed by multiple clone detection tools have a higher confidence to be real clones~\cite{cr2016ssbse}. By using this agreement-based approach, we could reduce the number of clone candidates for manual investigation by paying more attention to the ones agreed by multiple tools. To find an agreement between two clone pairs reported by two different tools, we used the clone pair matching metric proposed by Bellon et al.~\cite{Bellon2007}. Two clone pairs which have a large enough number of overlapping clone lines can be categorised as either a good-match or an ok-match pair (denoted as \textit{good} and \textit{ok} respectively) with a confident value between 0 and 1. A \textit{good} clone pair has stronger agreement than an \textit{ok} pair. We follow the original definitions of \textit{good}- and \textit{ok}-match introduced in Bellon's paper~\cite{Bellon2007}.

\vspace{0.5ex}
Given that a clone pair \textit{CP} is formed by two clone fragments \textit{CF$_1$} and \textit{CF$_2$} with a pre-defined similarity threshold \textit{t}, i.e.~\textit{CP} = (\textit{CF$_1$}, \textit{CF$_2$}, \textit{t}), we can define \textit{overlap} and \textit{contained} value of two clone pairs as 
%\squeezeup
\begin{multline}
	overlap(\textrm{\textit{CP}}_1, \textrm{\textit{CP}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1) \cup lines(\textrm{\textit{CF}}_2)|} 
\end{multline}
%\squeezeup
\begin{multline}
	contained(\textrm{\textit{CP}}_1, \textrm{\textit{CP}}_2) = \frac{|lines(\textrm{\textit{CF}}_1) \cap lines(\textrm{\textit{CF}}_2)|}{|lines(\textrm{\textit{CF}}_1)|}. 
\end{multline}

\noindent\textit{good-value} of two clone pairs is then defined as
%\squeezeup
\begin{align*}
	good(\textrm{\textit{CP}}_1, \textrm{\textit{CP}}_2) = min(overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1), \\ overlap(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2)).
\end{align*}

\noindent\textit{ok-value} is defined as
%\squeezeup
\begin{align*}
	ok(\textrm{\textit{CP}}_1,\textrm{\textit{CP}}_2) = min(max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1), \\ contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_1,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_1)),
	\\ max(contained(\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2), \\contained(\textrm{\textit{CP}}_2.\textrm{\textit{CF}}_2,\textrm{\textit{CP}}_1.\textrm{\textit{CF}}_2))).
\end{align*}

Two clone pairs $\textrm{\textit{CP}}_1$ and $\textrm{\textit{CP}}_2$ are called a \textit{\textit{good}(p)} iff, for threshold $p \in [0,1]$ holds 
%\squeezeup
\begin{equation}
good(\textrm{\textit{CP}}_1,\textrm{\textit{CP}}_2) \geq p.
\end{equation}

Similarly for an \textit{ok-match(p)} pair
%\squeezeup
\begin{equation}
ok(\textrm{\textit{CP}}_1,\textrm{\textit{CP}}_2) \geq p.
\end{equation}

Using this good and ok-match criteria with a predefined threshold \textit{p}, we can prune the 315 million candidate clone pairs for manual investigation. \textit{good} pairs are the ones with the highest confidence, followed by \textit{ok} pairs, and followed by clone pairs without agreement. We call the 
\textit{good} and \textit{ok} pairs \textbf{agreed} clone pairs and the the clone pairs reported by Simian and NiCad without agreement \textbf{disagreed} clone pairs.

\subsection{Clone Detector's Parameter Tuning}
We were aware of effects of configurations to clone detection results and the importance of searching for optimised configurations in empirical clone studies~\cite{Wang2014,cr2016ssbse,Ragkhitwetsagul2016,Svajlenko2014}. However, considering the massive size of the two datasets and search space of at least 15 Simian and 5 NiCad parameters, we were hindered from searching for the best configurations of the tools. Thus, we decided to configure Simian and NiCad using two established configurations: 1) the tools' default configurations chosen by the tools' creators (denoted as \textit{D}), and 2) the discovered configurations for Bellon's Java projects from \textit{EvaClone}, a study of optimising clone detectors' configurations based on clone agreement, by Wang et al.~\cite{Wang2013} (denoted by \textit{E}). The details of the two configurations are described in \Cref{t:param_tuning}. Having two clone detectors, Simian (denoted as \textit{S}) and NiCad (denoted as \textit{N}), with two chosen configurations each, we looked for Bellon's good and ok-match in four possible pair-wise combinations: $S_{D}N_{D}$, $S_{D}N_{E}$, $S_{E}N_{D}$, and $S_{E}N_{E}$.

\begin{table}
	\centering
	\caption{Configurations of Simian and NiCad}
	\label{t:param_tuning}
%	\resizebox{\columnwidth}{!}{%
	\small
	\begin{tabular}{p{0.7cm}|p{2.4cm}|p{3.5cm}}
		\hline 
		Tool & Default (\textit{D}) & EvaClone (\textit{E}) \\
		\hline
		Simian\newline (\textit{S}) &  threshold=6,\newline ignoreStringCase, \newline ignoreCharacterCase, \newline ignoreModifiers & threshold=5, ignoreIdentifiers, \newline ignoreIdentifierCase, \newline ignoreStrings, \newline ignoreCharacters,\newline ignoreSubtypeNames, \newline balanceSquareBrackets \\ 
		\hline 
		NiCad\newline (\textit{N}) & Blocks, UPI=0.30,\newline MinLine=10,\newline MaxLine=1000 & Blocks, UPI=0.20, \newline MinLine=5, MaxLine=604, \newline blind renaming,\newline literal abstraction \\
		\hline
	\end{tabular} %
%}
\end{table}

%\subsection{Manual Investigation}
%We did not have an oracle of clones
%between the two data sets so we rely on a manual investigation to
%validate the reported clone candidates. However, the large amount of
%generated clone pairs hindered us from looking at all of them.
%A random sampling of clones could have been employed, however, the 
%risk of returning false positives non-meaningful clones is increased.
%Instead of extracting a random sample, we selected clone pair
%candidates by relying on \emph{agreement} of clone detectors. If a
%clone pair was reported by multiple tools, we had a higher
%confidence that it was a real clone pair. To achieve this, clone pairs
%from the two clone detectors were pair-wise matched to find agreements
%using Bellon's clone overlapping criteria~\cite{Bellon2007}. This step
%generated 
%\textbf{agreed clone pairs}. They were pairs with high confidence to be true clones since
%they obtained agreement from both tools. Then, pairs reported by
%Simian and NiCad that did not find agreement were categorised as
%\textbf{disagreed clone pairs}. The disagreed clone pairs were clones
%with less confidence than the agreed ones. 
%Finally, agreed and
%disagreed clone pairs have been looked at manually by two of the
%authors.

%In the manual inspection process, we classified clones into seven
%categories according to information observed from code comments and
%text in the Stack Overflow answer. This process took approximately a
%month until we successfully investigated 3,632 clone pairs. Clone
%pairs that were classified as false clones, boiler-plate code, or
%IDE-generated are not very interesting and were ignore in the analysis
%of outdated code and licensing violation. We compared the licensing
%information of the 668 remaining clone pairs for possibilities of software
%licensing violations. Moreover, we looked at the history starting at
%the date of the answers to today from its git repository. With this
%method, we discovered \textbf{outdated code}, i.e.~code that changes
%were made to them after they have been copied to Stack
%Overflow. 
%hence resulting in outdated clones on Stack Overflow.

\section{Results and Discussion}

We followed the experimental framework and detected clones between Stack Overflow and Qualitas corpus using the two selected clone detectors. To answer RQ1, we computed statistics on clones discovered by the tools. To answer RQ2, we manually investigated online clone pair candidates and categorised them using our patterns of online code cloning. The patterns were derived from Kapser et al.~\cite{Kapser2003} and augmented by common patterns found in our data set. For RQ3, we looked at the true positive clone pairs and checked if they are still up-to-date. Similarly, for RQ4, we also looked at the license of each clone and checked for a possibility of licensing violation.

\subsection{RQ1: Online Code Clones}

\begin{table}
	\centering
	\caption{Statistics of online clones reported by Simian (\textit{S}) and NiCad (\textit{N}) with default (\textit{D}) and EvaClone (\textit{E}) configurations}
	\label{tab:raw_stats}
	\small
	\resizebox{\columnwidth}{!}{%
		\begin{tabular}{l|r|r|r|r}
			\hline
			Stats & \multicolumn{1}{c|}{$S_D$} & \multicolumn{1}{c|}{$S_E$} & \multicolumn{1}{c|}{$N_D$} & \multicolumn{1}{c}{$N_E$} \\
			\hline
			Stack Overflow Snippets & 1,086 & 1,531 & 1,392 & 12,886 \\
			No. of Qualitas projects & 92 & 111 & 82 & 88 \\
			Total C$_{\textrm{\textit{pairs}}}$ & 67,570 & 63,635,844 & 632,855 & 251,449,849 \\
%			Avg.~C$_{\textrm{\textit{pairs}}}$ & 62 & 41,565 & 455 & 19,356 \\
			Avg.~C$_{\textrm{\textit{size}}}$ & 9.11 & 5.95 & 10.85 & 6.28 \\
			Avg.~C$_{\textrm{\textit{\%}}}$ & 38\% & 50\% & 30\% & 33\% \\
			\hline
		\end{tabular} %
	}
\end{table}

\begin{figure*}
		\centering
		\begin{minipage}{.4\textwidth}
			\centering
			\includegraphics[width=\linewidth]{hist_sd}
			\caption*{$S_D$ clones}
		\end{minipage}%
		\hspace{5ex}
		\begin{minipage}{0.4\textwidth}
			\centering
			\includegraphics[width=\linewidth]{hist_nd}
			\caption*{$N_D$ clones}
		\end{minipage}
		\\
		\begin{minipage}{.4\textwidth}
			\centering
			\includegraphics[width=\linewidth]{hist_se}
			\caption*{$S_E$ clones}
		\end{minipage}%
		\hspace{5ex}
		\begin{minipage}{0.4\textwidth}
			\centering
			\includegraphics[width=\linewidth]{hist_ne}
			\caption*{$N_E$ clones}
		\end{minipage}
	\caption{\FIXME{Useful?}~Distribution of online clones between Stack Overflow snippets and Qualitas projects}
	\label{fig:histogram}
\end{figure*}

The clone statistics obtained from running Simian and NiCad are presented in \Cref{tab:raw_stats}. Simian with the default configuration ($S_D$) reports clones in 1,086 snippets, approximately 1\% of the 144,064 Stack Overflow snippets, associated with 92 Qualitas projects. Simian with the EvaClone configuration ($S_E$) reports 1,531 snippets containing clones associated with 111 Qualitas projects. NiCad with the default configuration ($N_D$) reports clones between 1,392 Stack Overflow code snippets and 82 Qualitas projects. NiCad with the EvaClone configuration ($N_E$) reports the largest number of 12,886 cloned snippets mainly due to its relaxed configurations. The $N_E$ snippets are associate with 88 Qualitas projects. We encountered NiCad failures with a few Qualitas projects.~$N_D$ could not detect clones in \textsf{hibernate} due to clustering errors. $N_E$ generated renaming errors for 5 projects including \textsf{vuze}, \textsf{hibernate}, \textsf{myfaces}, \textsf{netbeans}, and \textsf{spring}. We have contacted the creator of NiCad regarding the issues. They identified these issues as problems of long file paths and TXL grammars which will be fixed in the next NiCad releases. \Cref{fig:histogram} illustrates the distribution of online code clones reported by $S_D$, $N_D$, $S_E$ and $N_E$ respectively. We can see that the distribution is right skewed. There are only a few Stack Overflow snippets and Qualitas projects that are outliers containing a larger number of clones compared to the others.

In terms of the number of clone pairs, $S_D$ reports 67,570 clone pairs while $N_D$ reports 632,855 clone pairs. $S_E$ and $N_E$ report a larger clone pair amount of 63,635,844 and 251,449,849 pairs respectively. This is expected since EvaClone configurations prefer recall~\cite{Wang2013}. 
%Using $S_D$, we found 62 clones in average originating from a single Stack Overflow snippet. 62 and 455 respectively. On the other hand, $S_E$ and $N_E$ have a much larger number of 63 and 251 million clone pairs per snippet on average. 
The average clone size reported by $S_D$ is 9.11 lines which is bigger than $S_E$ (5.95 lines). Similarly, $N_D$ has an average clone size of 10.85 lines which is bigger than 6.28 lines reported by $N_E$. We can see from the statistics that EvaClone tunes the tools to report smaller clones. For the cloned Stack Overflow snippets, the average ratio of cloned code for $S_D$, $S_E$, $N_D$, and $N_E$ is 38\%, 50\%, 30\%, and 33\%  respectively. 
%A quick manual check of Simian's clone report revealed that there were 11 problematic snippets. These 11 snippets triggered Simian to generate large clone clusters containing a huge number of false clones from array initialisation. Hence, their clones were removed from further analyses.

\textbf{To answer RQ1, with the four configurations of Simian and NiCad, we found online code clones in 1,086 ($S_D$), 1,531 ($S_E$), 1,392 ($N_D$) and 12,886 ($N_E$) Stack Overflow code snippets. They are associated with 92, 111, 82 and 88 Qualitas projects respectively.}

\subsection{RQ2: Patterns of Online Code Cloning}

We rely on a manual study to qualitatively investigate the patterns of online code cloning. To achieve this, we select a subset of the agreed clones using Bellon's criteria and a subset of the disagreed clones using clone filtering. We explain them in more detail below.

\subsubsection{Agreed clone pairs}

As displayed in \Cref{t_agreed_good_clone_pairs}, after we applied clone agreement filtering using Bellon's criteria, the number of \textit{good} clone pairs is 2,308 while the number of \textit{ok} clone pairs is 31,729. %The number of disagreed clone pairs ($S_D$, $N_D$) after filtering is 17,801 and 181 for Simian and NiCad respectively. The total number of clone candidates is 32,039 pairs. 
Agreed clone pairs are clone pairs that pass Bellon's \textit{good} or \textit{ok} criteria, and therefore are selected for manual classification. Similar to the original study, we select a threshold \textit{p} of 0.7 for both \textit{good} and \textit{ok} pairs~\cite{Bellon2007}. %The number of projects processed by the clone detection tools are listed in \Cref{tab:projects_missing}. 
Due to NiCad's failures, we do not have NiCad clones from five Qualitas projects in the agreed clone pairs. Simian clones of the same projects also disappeared from the agreed clone pairs since they could not find matches.

\begin{table}
	\centering
	\caption{Distribution of the agreed clone pairs}
	\label{t_agreed_good_clone_pairs}
	\small
%	\resizebox{\columnwidth}{!}{%
	\begin{tabular}{l|r|r|r|r|r}
		\hline
		Set & $S_DN_D$ & $S_DN_E$ & $S_EN_D$ & $S_EN_E$ & Unique \\
		\hline
		\textit{good} & 18 & 26 & 10 & 2,267 & 2,308 \\
		\textit{ok} & 11,467 & 993 & 79 & 20,021 & 31,729  \\
		\hline
	\end{tabular} %
%}
\end{table}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{good-ok_pairs-crop}
	\caption{Distribution of \textit{good} (top) and \textit{ok} pairs (bottom) reported by the 4 configurations of Simian and NiCad. The shaded areas are overlapping pairs.}
	\label{fig:good-ok-pairs}
\end{figure}

The distributions of \textit{good} clone pairs between four combinations of \textit{D} and \textit{E} configurations are listed in \Cref{t_agreed_good_clone_pairs}. %and depicted visually in \Cref{fig:good-ok-pairs}. 
There are 2,321 (2,308 unique) \textit{good} pairs consisting of 18 pairs from $S_DN_D$, 26 pairs from $S_DN_E$, 10 pairs from $S_EN_D$, and 2,267 pairs from $S_EN_E$. There are 32,560 (31,729 unique) \textit{ok} pairs\footnote{This excludes the subsumed 2,308 \textit{good} pairs. According to the definition, \textit{ok} pairs always subsume \textit{good} pairs.} We obtained 11,467; 993; 79; and 20,021 pairs from $S_DN_D$, $S_DN_E$, $S_EN_D$, and $S_EN_E$ respectively. Between the four configuration sets, there is considerable amount of clone pairs shared between two adjacent sets (as depicted in \Cref{fig:good-ok-pairs}), but there is no clone pair that is agreed by all four combinations.

\subsubsection{Disagreed clone pairs}

\begin{table}
	\centering
	\caption{Disagreed clone pairs before and after filtering.}
	\label{tab:online_clone_pairs}
	\small
	%	\resizebox{\columnwidth}{!}{%
	\begin{tabular}{l|r|c|c|c|c|r}
		\hline
		\multirow{2}{*}{Set} & \multirow{2}{*}{Before Filtering} & \multicolumn{4}{c|}{Filters*} & \multirow{2}{*}{After Filtering} \\ \cline{3-6}
		& & \textit{L} & \textit{S} & \textit{R} & \textit{G} & \\
		\hline 
		%		\multirow{1}{*}{\textit{good}} & 2,308 & & & & & 2,308 \\
		%		\multirow{1}{*}{\textit{ok}} & 11,749 & & & & & 11,749 \\
		%\multirow{1}{*}{\textit{agreed clone pairs}}      & 29,099 \\
		\multirow{1}{*}{$S_D$} & 67,570 & \checkmark & & \checkmark & \checkmark & 17,801 \\
		\multirow{1}{*}{$N_D$} & 632,855 & & \checkmark & \checkmark & \checkmark & 88 \\ 
		%\multirow{1}{*}{\textit{disagreed clone pairs}}    & 1,039 \\
%		\hline
%		Total  & 6,400,425 & \multicolumn{4}{c|}{ } & 17,169 \\ 
		\hline
		\multicolumn{7}{l}{*\textit{L}=line $\geq 10$, \textit{S}=sim $\geq 85\%$, \textit{R}=reg.~expressions, \textit{G}=\textit{good}/\textit{ok} pairs.}  \\
	\end{tabular} %
	%}
\end{table}

The disagreed clone pairs are clone pairs that are reported by a single tool, either Simian or NiCad, and do not have agreement with the other tool. The disagreement can be from a misalignment of clone lines or totally non-overlapping clones. The disagreed pairs from Simian cover clone pairs in the projects with NiCad's errors (1 projects from $N_D$ and 5 from $N_E$). %that are automatically missing from the agreed clone pairs. 
With the four configuration combinations, we decided to investigate only $S_D$, and $N_D$, and dropped $S_E$ and $N_E$ due to their enormous amount of clone pairs (60 and 250 millions respectively). Moreover, $S_E$ and $N_E$ contain a large number of false positives due to the recall preference of their EvaClone configurations.

Even choosing only the default configurations, the number of clone pair candidates are still very large for manual inspection (67,570 for $S_D$ and 632,855 pairs for $N_D$). We applied four filters: \textbf{clone line, regular expression, similarity threshold, and good/ok pairs}. For the clone line filter, we raised the minimum clone size to 10 line since larger clones are more interesting, while smaller ones tend to be false clones~\cite{Saini2016}. Since the 10--line threshold is already the default configuration for NiCad, this filter is applied to Simian only (Simian's default configurations has a minimum clone size of 6 lines).~The regular expression filter removes boiler-plate clones of \verb|hashCode()| methods and getters/setters. The third filter, similarity threshold, applies only to NiCad clone pairs since Simian does not provide this similarity configuration. The last filter, good/ok pairs, removes the clone pairs that have already existed in the agreed clone sets. 
% Moreover, the study of uniqueness of source code by Gabel and Su~\cite{Gabel2010} suggests that source code are unique when they are larger than seven lines.

\begin{table}
	\centering
	\caption{No.~of online clones candidates for the classification}
	\label{tab:online_clone_classification_results}
	\small
	%	\resizebox{\columnwidth}{!}{%
	\begin{tabular}{l|r|r|r}
		\hline
		\multirow{2}{*}{Set} & \multirow{2}{*}{Candidates} & \multicolumn{2}{c}{Classification} \\ \cline{3-4}
		& & Auto & Manual \\
		\hline 
		\multirow{1}{*}{\textit{good}} & 2,308 & 10 & 2,298 \\
		\multirow{1}{*}{\textit{ok}} & 11,749 & 11,494 & 255 \\
		%\multirow{1}{*}{\textit{agreed clone pairs}}      & 29,099 \\
		\multirow{1}{*}{$S_D$} & 17,801 & 16,800 & 1,001 \\
		\multirow{1}{*}{$N_D$} & 88 & 10 & 78 \\ 
		%\multirow{1}{*}{\textit{disagreed clone pairs}}    & 1,039 \\
		\hline
		Total  & 31,946 & 28,314 & 3,632 \\ 
		\hline
	\end{tabular} %
	%}
\end{table}

As shown in \Cref{tab:online_clone_pairs}, for $S_D$ we filtered the results using clone size, regular expressions, and gook/ok pairs. The three filters reduced the number of $S_D$ clones to 17,801 pairs. 
%For $N_D$, 10-line threshold is already a minimum NiCad's clone size. However, NiCad still reported much more clone pairs than Simian since it can detect type-3 clones. Thus, besides regular expressions and good/ok pair filters, 
We filtered $N_D$ clones by raising the similarity threshold. We increase NiCad's similarity threshold from 70\% to 85\% (by adjusting NiCad's $\mathrm{UPI}$ parameter from 0.3 to 0.15). We tried varying the UPI from 70\% (default -- 632,855 pairs), 80\% (25,003 pairs), 85\% (185 pairs) and 90\% (83 pairs) and found that 85\% similarity provided the best balance in terms of the amount of reported clones and feasibility for manual investigation. Using this filter along with the regular expression and the good/ok pair filter, we reduced the size of $N_D$ clones to 88 pairs.

%In our clone classification process, we used a script to automatically classify {\small\texttt{equals()}} clone pairs which  classifying 28,444 pairs out of 31,946 (more than half of the {\small\texttt{equals()}} pairs came from \textit{ok} and $S_D$ pairs). Finally, there are 3,502 pairs remaining which we manually classified them. The classification summary is shown in \Cref{tab:online_clone_classification_results}.

%After classification, we found 668 non-trivial online code clones pairs between Stack Overflow and Qualitas or external sources.}

\begin{table}
	\centering
	\caption{The seven patterns of online code cloning}
	\label{tab:classification_scheme}
	\resizebox{\columnwidth}{!}{%
		\begin{tabular}{c|p{8.4cm}}
			\hline 
			Patt. & Descriptions \\ 
			\hline 
			QS & Cloned from Qualitas project to Stack Overflow (Q $\rightarrow$ S). \\ 
			\hline 
			SQ &Cloned from Stack Overflow to Qualitas project (S $\rightarrow$ Q). \\ 
			\hline 
			EX & Cloned from an external source (X $\rightarrow$ S $\wedge$ X $\rightarrow$ Q). \\
			\hline 
			UD & Cloned from each other or from an external source \textit{X} outside the project (unknown) (S $\leftrightarrow$ Q $\vee$ (X $\rightarrow$ S $\wedge$ X $\rightarrow$ Q)). \\ 
			\hline 
			BP & Boiler-plate or IDE auto-generated \\ 
			\hline 
			IN & Inheritance, interface implementation  \\ 
			\hline 
			AC & Accidental similarity, false clones \\ 
			\hline 
		\end{tabular}  %
	}
\end{table}

\subsubsection{The Seven Patterns of Online Code Cloning}

We are interested in the analysis of the patterns of online code clones. We studied the 8 patterns of cloning from Kapser et al.~\cite{Kapser2006,Kapser2008} and performed a small study to evaluate its applicability to our study. Using the $S_D$ clone report, snippets are ranked according to 1) clone pair frequency, 2) popularity (i.e.~number of associated Qualitas projects), 3) clone size in SLOC, and 4) ratio of cloned code per snippet. We selected these four criteria so that we could cover the online clones from various aspects. We selected the top 10 from each ranking and obtained 34 unique Stack Overflow snippets. The snippets were associated with 697 clone pairs. Using Kapser's cloning patterns, the clone pairs were categorised into either Customisation or Templating.

The benefit of the study is two-folded. We evaluated Kapser's cloning patterns and also observed the common cloning patterns from the 697 clone pairs. \textbf{Clearly Kapser's cloning patterns are too broad for our study and a more suitable and fine-grained classification scheme is needed.} After the study, we adopted one of Kapser's cloning patterns, boiler-plate code and API/library protocol, and defined 6 new cloning patterns . The seven patterns include QS, SQ, EX, UD, BP, IN, and AC as presented in \Cref{tab:classification_scheme}. Pattern QS (\textbf{Q}ualitas to \textbf{S}tack Overflow) represents clones that have evidence to be copied from Qualitas to Stack Overflow. The evidence of copying can be found in comments in the Qualitas source code or in the Stack Overflow post's contents. Pattern SQ (\textbf{S}tack Overflow to \textbf{Q}ualitas) is cloning, with evidence, in the opposite direction to QS. Pattern EX (\textbf{Ex}ternal Sources) is cloning that has evidence of copying from a single or multiple external sources.  Pattern UD (\textbf{U}nknown \textbf{D}irection) is cloning that creates identical or highly similar clones between Qualitas and Stack Overflow but where we could not find any attribution of copying. %The UD clones are also unique enough to not be a boiler-plate code. 
Pattern BP (\textbf{B}oiler-\textbf{P}late) represents clones containing boiler-plate code of {\small\verb|equals()|} methods, getters/setters, or IDE-generated code such as GUI components. Pattern IN (\textbf{In}heritance/Interface) is cloning by inheritance of the same super class or implementation of the same interface. These two activities usually result in similar overriding methods. The last pattern, AC (\textbf{A}ccidental \textbf{C}lones), represents accidentally similar clone pairs. These are mainly false positive clones from the clone detectors such as similar {\small\texttt{try-catch}} statements.

%With the online cloning patterns, we could calculate the number of true and false positives. %We are looking for unique and meaningful clones in this study. Thus,

\subsubsection{Classification of Clone Candidates}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{classification_process}
	\caption{Online code clone classification process}
	\label{fig:classification_process}
\end{figure}

The online clone classification follows the steps depicted in \Cref{fig:classification_process}. First, we look at a pair of clone fragments to see their similarity. If they are accidentally similar clones after code normalisation or false positive clones from the clone detection tools, we classify the pair into AC. If the two fragments are boiler-plate code, the pair is classified into BP. If they implement the same interface or inherited the same class and share similar overriding methods, the pair is classified into IN. If the pair is not BP, IN, or AC, we start a detailed investigation. We check the corresponding Stack Overflow post, read through it carefully and look for any evidence mentioning code copying. If evidence of copying has been found from a Qualitas project, the pair is classified in QS. In several occasions, we used extra information such as questions' contents, name of posters, and tags to gain better understanding. On the other hand, if the source code from the Qualitas project mentions copying from Stack Overflow, the pair is classified into SQ. If there is evidence of copying from an external source instead of Qualitas projects, the pair is classified into EX. Lastly, if there is no evidence of copying in any direction but the clone fragments are highly similar, we classify them into UD.

We had in total 31,946 online clone pair candidates for classifications. 2,308 and 11,749 were \textit{good} and \textit{ok} clone pairs respectively. 17,801 and 88 clone pairs are from $S_D$ and $N_D$ after the filtering. Please note that we ignored 20,021 $S_EN_E$ \textit{ok} clone pairs in the classification due to their large number of clones and the very small average clone sizes of 4.79 and 5.33 lines (see \Cref{tab:raw_stats}). They had a high probability to be trivial clones.

Before performing manual classification, we used regular expressions to automatically classify {\small\texttt{equals()}} and the remaining getters/setters. As shown in \Cref{tab:online_clone_classification_results}, the automatic classifier helped us classified and captured 28,314 clone pairs into pattern BP.

There were 3,632 pairs left for the manual check. The first author who has been working on clone detection research for two years took the role of the main investigator performing a manual classification of the 3,632 clone pairs. By consulting the cloning patterns, the main investigator went through each clone pair candidate, looked at the clones, and chose the most appropriate cloning pattern for the pair. %A relevant and useful observation was also recorded for each clone pair.

To mitigate the human errors of the manual classification, the second author took the role of the validating investigator, performing a validation on 10\% (363 pairs) of the clone pairs judged by the main investigator. After the validation, the classifications from the two investigators were compared and 64 conflicts were found. The two investigators discussed and resolved 61 of them while the 3 conflicts were resolved by involving the third author. The classification results are shown in \Cref{tab:classification_good_o}. %We consider the clone pairs classified into pattern QS, SQ, UD, and EX as true positives. On the other hand, clones in pattern BP, IN, and AC are classified as false positives since they are trivial and not very interesting ones.
\textbf{To answer RQ2, the classification results are explained and discussed below.}

\textbf{QS: Qualitas $\rightarrow$ Stack Overflow.} We found 112 online clone pairs with evidences of cloning from Qualitas projects to Stack Overflow. They include 3, 14, 86, and 9 pairs from \textit{good}, \textit{ok}, $S_D$, and $N_D$ set respectively. Most of them have a statement in the Stack Overflow post saying that the code is ``copied'' or ``modified'' from a specific file or class in a Qualitas project. The most cloned projects are \textsf{hibernate} and \textsf{spring} with 20 clone pairs, followed by \textsf{eclipse} (15 pairs), and \textsf{hadoop} (12 pairs). The clones are used as examples and are very similar to their original Qualitas code with limited modifications.

\textbf{SQ: Stack Overflow $\rightarrow$ Qualitas.} We did not find any clone pairs with evidence of cloning from Stack Overflow to Qualitas projects.

\textbf{EX: External Sources.} We found 197 online clone pairs with evidences of cloning from external sources to Stack Overflow. They include 4, 41, 146, and 6 clone pairs from \textit{good}, \textit{ok}, $S_D$, and $N_D$ respectively.

These findings complement a study of clones between software projects~\cite{Svajlenko2014}. We found that the code cloning can also happen among different sources on the Internet just like software projects. There are 30 clone pairs that are originated from programming websites. For example, we found a clone of code snippet to convert numbers into words which is copied from www.rgagnon.com/javadetails/java-0426.html. Both the Stack Overflow and the \textsf{compiere} project source code contain an attribution to the original source. Another example is a code to generate graphical \textit{Perlin noise}. It is copied from http://mrl.nyu.edu/\textasciitilde perlin/noise/. The code is used in Stack Overflow and \textsf{aoi} project with attribution. %The original code of the 15 clone pairs do not have licensing information so these clones do not violate any license. 
There are other EX clone pairs are from third-party libraries such as \textsf{zxing}, \textsf{jasper}, \textsf{java.util}, \textsf{javax.servlet}, and the \textsf{bouncycastle} cryptography project.

\textbf{UD: Unknown Direction.} We found 382 online clone pairs with no evidence of cloning between Qualitas and Stack Overflow but with a high code similarity. They include 11, 42, 294, and 35 pairs from \textit{good}, \textit{ok}, $S_D$, and $N_D$ set respectively. The most cloned projects is \textsf{netbeans} with 95 clone pairs. Most of the clones are a large chunk of code handling GUI components. Although these GUI clones might be auto-generated by IDEs, we did not find any evidence. The second most cloned project is \textsf{jung2} (62 pairs), a network and graph framework, followed by \textsf{eclipse-SDK} (37 pairs). %There are several Stack Overflow snippets that are identical to code in \textsf{jung2}.

\textbf{BP: Boiler-Plate.} There were a large amount of boiler-plate clone pairs found in this study. Although we have removed some of the boiler-plate code with filters before the analysis, we still observed 28,749 clone pairs. The BP clone pairs account for 89.7\% of all clone pairs we classified. The majority of them are {\small{\texttt{equals()}}}.

\textbf{IN: Inheritance/interface.} There were 111 clone pairs found to be similar because they implement the same interface or inheriting from the same class. An example is the two implementations of {\small{\texttt{MouseListener}}} that share similar {\small\texttt{@Override}} methods of {\small\texttt{mousePressed}}, and {\small\texttt{mouseReleased}}.

\textbf{AC: Accidental Clones.} There were 2,418 accidental clone pairs. Mainly, they are false positive clones caused by code normalisation. %However, the original source code were still different enough when we manually classified them. 
Other clone instances include {\small\texttt{finally}} or {\small\texttt{try-catch}} clauses that were accidentally the same due to their very small sizes, and similar {\small\texttt{switch-case}} statements.

\begin{table}
	\centering
	\caption{Classifications of online clone pairs.}
	\label{tab:classification_good_o}
	\small
	\resizebox{\columnwidth}{!}{%
		\begin{tabular}{l|r|r|r|r|r|r|r|r}
			\hline
			Set & QS & SQ & UD & EX & BP & IN & AC & Total \\ 
			\hline
			\multirow{1}{*}{\textit{good}}  & 3 & 0 & 11 & 4 & 111  & 7 & 2,172 & 2,308 \\
			\multirow{1}{*}{\textit{ok}}  & 14 & 0 & 20 & 41 & 11,547 & 22 & 105 & 11,749 \\
			\multirow{1}{*}{$S_D$} & 86 & 0 & 293 & 146 & 17,074 & 74 & 128 & 17,801 \\
			\multirow{1}{*}{$N_D$} & 9  & 0 & 35 & 6 & 17 & 8 & 13 & 88 \\ 
			\hline
			Total & 112 & 0 & 359 & 197 & 28,749 & 111 & 2,418 & 31,946 \\ 
			\hline
		\end{tabular} 
	}
\end{table}

%For 17,801 $S_D$ clone pairs, they contained a large volume of boiler-plate code. Our regular expressions could automatically classified 16,800 of them into BP. The investigator checked the 1,001 remaining pairs manually. As a result, we found 86 QS, 294 UD, 145 EX, 17,074 BP, 74 IN and 128 AC pairs. For 88 $N_D$ clone pairs, 10 of them were automatically classified into BP and 78 were manually classified. We found 9 QS, 35 UD, 6 EX, 17 BP, 8 IN, and 13 AC respectively.

%\textbf{We answer RQ2 by using our patterns of online code cloning to classified 31,946 clone pairs. We found that a large amount of the clone pairs are boiler-plate code (28,642), clones from inheritance/interface (109), or accidental clones (2,525). Nevertheless, there were 670 pairs that were intentionally created. 110 pairs had strong evidences to be cloned from Qualitas to Stack Overflow. 382 were highly-confident clones but without evidence of cloning direction. 178 pairs were found to be cloned from external sources outside Qualitas projects.} %We did not find any clone pairs in the direction from Stack Overflow to Qualitas.}

\subsection{RQ3: Outdated Online Code Clones}

%In this study, we are interested in effects of having code clones between open source software systems and Stack Overflow. From the manual investigation, we found 523 true positive online clone pairs. With this set of true clones, we investigated further and found that there are two potential issues, outdated code and software licensing violation.

%\subsubsection{Outdated Clones}
Outdated code occurs when a piece of code has been copied from its origin to another location and later the original has been updated~\cite{Xia2014}. Usually code clone detection is used to locate clone instances and update them to match with the originals~\cite{Bellon2007}. However, online code clones are more difficult to detect than in regular software projects due to its large search space and a mix of natural and programming languages combined in the same post. 
%Moreover, Stack Overflow is not a software. There is no test suite to run and check whether the validity of the code snippets. To the best of our knowledge, this work is the first to address the problem of outdated code on Stack Overflow.

To search for outdated online code clones, we focused on the 112 QS clone pairs that were cloned from Qualitas to Stack Overflow and compared them with their latest versions. We downloaded the latest version of the Qualitas projects from their repositories on 26 September 2016. For each QS online clone pair, we use the clone from Qualitas a a proxy. We searched for its latest version by the file name and located the cloned region in the file based on the method name. 
%For Simian clones, where clones may span across several methods, we scanned upwards until we find a method name ahead of the clone region. 
We then compared the Stack Overflow snippet to its latest version line-by-line to find if any change has been made to the source code. We also made sure that the changes did not come from the modifications made to the Stack Overflow snippets by the posters but from the updates in the projects themselves. When we found inconsistent lines between the two versions, we used {\small\texttt{git blame}} to see who modified those lines of code and the timestamps.

 %Some of the clone pairs overlap with each other so we group them in the same row. 
We discovered 64 outdated online clone pairs out of 112 pairs. As shown in \Cref{fig:outdated}, \textsf{spring} and \textsf{hibernate} have the maximum number of 20 outdated pairs, followed by 15 from \textsf{eclipse}, and 12 from \textsf{hadoop}. Besides the example of outdated code in the \textsf{hadoop}'s {\small{\texttt{WritableComparator.java}}} shown in \Cref{fig:before-after}, we also found a few outdated code elements which contained heavy modifications. For example, the code snippet in Stack Overflow post 23520731 is a copy of {\small{\texttt{SchemaUpdate.java}}} in \textsf{hibernate}. The code has been heavily modified on 5 February 2016. In the 64 outdate pairs, we found 5``dead'' snippets. These snippets cannot be located in the latest version of the projects. For example, the snippet in Stack Overflow post 3758110, a copy of {\small{\texttt{DefaultAnnotationHandlerMapping.java}}} in \textsf{spring}, was deleted in the commit {\small{\texttt{02a4473c62d8240837bec297f0a1f3cb67ef8a7b}}} by Chris Beams on 20th January 2012, two years after it was posted. %The latest version of Spring does not contain the file anymore. 
\Cref{tab:stale_code_details} shows examples of the outdated online clones on Stack Overflow. The table displays information of the clones from both Stack Overflow and Qualitas side including the dates. We summarise the changes that make the clones outdated into three types, modified/added/deleted statements (\textit{S}), file deletion (\textit{D}), and method rewriting (\textit{R}), along with the date of change.

The outdated online code clones cause problems ranging from uncompilable code (due to modifications and different API usage in the outdated code) to introducing vulnerabilities to software~\cite{Xia2014}. An outdated code with a subtle change (e.g.~\Cref{fig:before-after}) may be copied and reused without awareness from developers. Although Stack Overflow has a voting mechanism that may mitigate this issue, the accepted answer is still attractive enough for naive developers to copy and reuse the code.

\textbf{For RQ3, our results show that more than half (64) of QS clone pairs on Stack Overflow are outdated. 60 pairs differ from their newest versions by modifications applied to variable names or method names, added or deleted statements, to a fully rewritten code with new method signatures. 4 pairs are dead snippets.} %They do not exist any more in the latest version.} %These outdated Stack Overflow online clone snippets are questionable for reuse.} %These changes did not reflect They might later find out that the copied code does not work any more due to different API versions. Even worse, they might also introduce bugs or vulnerabilities into their software.  }
%Usually when developers reuse a code snippet from a Stack Overflow post and find that it does not work nor compatible with their environment. They can cast a down vote to that answer resulting in low votes for the answer (which might be outdated). However, if the answer is marked as accepted by the person who asks the question, the check mark symbol attached to the answer is still attractive to naive developers who are ignorant.

%\begin{figure*}
%	\begin{lstlisting}
%          /* Code in Stack Overflow #23520731 */             /* SchemaUpdate.java (2016-09-26) */
%          public void execute (Target target) {              public void execute(EnumSet<TargetType> targetTypes, 
%            LOG.runningHbm2ddlSchemaUpdate();                                Metadata metadata, ServiceRegistry serviceRegistry) {
%            Connection connection = null;                      if ( targetTypes.isEmpty() ) {
%            Statement stmt = null;                               LOG.debug(""Skipping SchemaExport as no targets were specified"");
%            Writer outputFileWriter = null;                      return;
%            exceptions.clear();                                }
%            try {                                              exceptions.clear();
%              DatabaseMetadata meta;                           LOG.runningHbm2ddlSchemaUpdate();
%              ...                                              ...
%	\end{lstlisting}
%	\caption{Outdated code snippet on Stack Overflow post 23520731. This code has been copied from {\small\texttt{SchemaUpdate.java}}. Its latest version in \textsf{hibernate} code base contains heavy modifications.}
%	\label{fig:hibernate_outdated_code}
%\end{figure*}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{freshness}
	\caption{Status of the 112 QS online clone pairs}
	\label{fig:outdated}
\end{figure}

%\begin{table*}
%	\centering
%	\caption{Examples of the 6 outdated QS online clones. Outdated clones are caused by modified/added/deleted statements (\textit{S}), file deletion (\textit{D}), and method rewriting (\textit{R}).}
%	\label{tab:stale_code_details}
%	\small
%	\resizebox{2\columnwidth}{!}{%
%	\begin{tabular}{r|l|l|l|l|p{5.2cm}|r|r|l|c|l}
%		\hline 
%		\multirow{2}{*}{No.} & \multicolumn{2}{c|}{Stack Overflow} & \multicolumn{6}{c|}{Qualitas} & \multicolumn{2}{c}{Changes} \\ \cline{2-11}
%		& Post & Date & Project & Ver. & File & Start & End & Date & Type & Date \\
%		\hline
%			1 & 18303692 & 18-Aug-13 & aspectj & 1.6.9  & aspectjtools/../Agent.java & 7 & 18 & 5-Mar-08 & \textit{S} & 8-Sep-15 \\
%			2 & 18303692 & 18-Aug-13 & aspectj & 1.6.9  & aspectjweaver/../Agent.java & 7 & 18 & 5-Mar-08 & \textit{S} & 8-Sep-15 \\
%			3 & 2513183 & 25-Mar-10 & eclipse-SDK & 4.3 & GenerateToStringAction.java & 113 & 166 & 5-Jun-13 & \textit{S} & 17-Mar-15 \\
%			4 & 2513183	& 25-Mar-10 & eclipse-SDK & 4.3 & GenerateToStringAction.java & 117 & 126 & 5-Jun-13 & \textit{S} & 17-Mar-15 \\
%			5 & 2513183	& 25-Mar-10 & eclipse-SDK & 4.3 & GenerateToStringAction.java & 143 & 165 & 5-Jun-13 & \textit{S} & 1-Mar-11 \\
%			6 & 2513183	& 25-Mar-10 & eclipse-SDK & 4.3 & GenerateToStringAction.java & 178 & 187 & 5-Jun-13 & \textit{S} & 5-Jun-13 \\
%			7 & 11861598 & 8-Aug-12 & eclipse-SDK & 4.3 & WizardDialog.java & 377 & 394 & 1-May-13 & \textit{S} & 21-Jun-13 \\
%			8 & 22315734 & 11-Mar-14 & hadoop & 1.0.0 & WritableComparator.java & 44 & 54 & 25-Aug-11 & \textit{S} & 20-Nov-14 \\
%			9 & 801987 & 29-Apr-09 & hadoop & 1.0.0 & StringUtils.java & 40 & 56 & 15-Dec-11 & \textit{S} & 4-Feb-13 \\
%			10 & 21702608 & 11-Feb-14 & hadoop & 1.0.0 & DBCountPageView.java & 275 & 287 & 15-Dec-11 & \textit{S} & 12-Jun-11 \\
%			11 & 21702608 & 11-Feb-14 & hadoop & 1.0.0 & DBCountPageView.java & 289 & 309 & 15-Dec-11 & \textit{S} & 12-Jun-11 \\
%			12 & 14845581 & 13-Feb-13 & hadoop & 1.0.0 & JobSubmissionFiles.java & 46 & 55 & 15-Dec-11 & \textit{S} & 25-Jun-12 \\
%			13 & 16180910 & 24-Apr-13 & hadoop & 1.0.0 & mapred/../LineRecordReader.java & 47 & 60 & 15-Dec-11 & \textit{R} & 25-Jul-11 \\
%			14 & 16180910 & 24-Apr-13 & hadoop & 1.0.0 & mapreduce/../LineRecordReader.java & 41 & 54 & 15-Dec-11 & \textit{R} & 25-Jul-11 \\
%			15 & 15168494 & 24-Feb-14 & hibernate & 4.2.2 & ConnectionProviderInitiator.java & 65 & 93 & 22-May-13 & \textit{S} & 26-Apr-13 \\
%			16 & 24924255 & 24-Jul-14 & hibernate & 4.2.2 & Example.java & 224 & 243 & 22-May-13 & \textit{S} & 23-Apr-13 \\
%			17 & 23520731 & 7-May-14 & hibernate & 4.2.2 & SchemaUpdate.java & 115 & 168 & 22-May-13 & \textit{S} & 5-Feb-16 \\
%			18 & 8257554 & 26-Nov-11 & hibernate & 4.2.2 & SettingsFactory.java & 244 & 255 & 22-May-13 & \textit{D} & 11-Mar-11 \\
%			19 & 23967852 & 31-Apr-14 & hibernate & 4.2.2 & SQLServer2005LimitHandler.java & 43 & 61 & 22-May-13 & \textit{S} & 11-May-16 \\
%			20 & 19298607 & 10-Oct-13 & hibernate & 4.2.2 & Oracle9iDialect.java & 23 & 32 & 22-May-13 & \textit{S} & 12-Apr-15 \\
%			21 & 10031897 & 5-Apr-12 & hibernate & 4.2.2 & AliasToBeanResultTransformer.java & 44 & 61 & 22-May-13 & \textit{S} & 4-Jun-15 \\ 
%			22 & 8037824 & 7-Nov-11 & jasperreports & 3.7.4 & JRVerifier.java & 982 & 998 & 31-May-10 & \textit{S} & 17-Apr-08 \\
%			23 & 8037824 & 7-Nov-11 & jasperreports & 3.7.4 & JRVerifier.java & 1221 & 1240 & 31-May-10 & \textit{D} & 20-May-11 \\
%			24 & 12936580 & 17-Oct-12 & jfreechart & 1.0.13 & AbstractXYItemRenderer.java & 532 & 569 & 20-Apr-09 & \textit{S} & 16-Jan-16 \\
%			25 & 16058183 & 17-Apr-13 & jfreechart & 1.0.13 & KeyToGroupMap.java & 18 & 30 & 20-Apr-09 & \textit{S} & 29-Jun-07 \\
%			26 & 21998949 & 24-Feb-14 & jfreechart & 1.0.13 & SpiderWebPlot.java & 502 & 520 & 20-Apr-09 & \textit{S} & 2-Jun-08 \\
%			27 & 21998949 & 24-Feb-14 & jfreechart & 1.0.13 & SpiderWebPlot.java & 522 & 536 & 20-Apr-09 & \textit{S} & 22-Nov-13 \\
%			28 & 6722760 & 17-Jul-11 & jgraph & 5.13.0.0 & HelloWorld.java (3 overlaps)  & 14 & 37 & 25-Sep-09 & \textit{R} & 13-Apr-14 \\
%			31 & 6025026 & 17-Jul-11 & jung2 & 2\_0\_1  & ShortestPathDemo.java & 106 & 117 & 20-Jun-08 & \textit{S} & 13-Apr-10 \\
%			32 & 6025026 & 17-Jul-11 & jung2 & 2\_0\_1  & ShortestPathDemo.java & 158 & 172 & 20-Jun-08 & \textit{S} & 29-Nov-15 \\
%			33 & 23586872 & 31-May-14 & junit & 4.11 & Assert.java & 33 & 52 & 11-Oct-12 & \textit{S} & 12-May-15 \\
%			34 & 7504040 & 29-Sep-11 & junit & 4.11 & ExternalResource.java (3 overlaps) & 4 & 23 & 11-Oct-12 & \textit{S} & 25-Jun-16 \\
%			37 & 8802082 & 10-Jan-12 & junit & 4.11 & ExpectException.java (2 overlaps) & 11 & 29 & 11-Oct-12 & \textit{S} & 25-May-14 \\
%			39 & 12593810 & 26-Sep-12 & poi & 3.6 & WorkbookFactory.java & 18 & 28 & 7-Dec-09 & \textit{R} & 26-Jun-13 \\
%			40 & 20913543 & 6-Jan-14 & spring & 3.0.5  & AutowireUtils.java & 32 & 42 & 20-Oct-10 & \textit{S} & 28-Oct-14 \\
%			41 & 18623736 & 4-Sep-13 & spring & 3.0.5  & CustomCollectionEditor.java & 33 & 71 & 20-Oct-10 & \textit{S} & 21-Nov-13 \\
%			42 & 22865824 & 4-Apr-14 & spring & 3.0.5  & JavaMailSenderImpl.java & 169 & 185 & 20-Oct-10 & \textit{S} & 6-Oct-14 \\
%			43 & 22865824 & 4-Apr-14 & spring & 3.0.5  & JavaMailSenderImpl.java & 186 & 197 & 20-Oct-10 & \textit{S} & 6-Oct-14 \\
%			44 & 20421869 & 6-Dec-13 & spring & 3.0.5  & ClassPathScanningCandidateComponent\newline Provider.java (2 overlaps) & 85 & 133 & 20-Oct-10 & \textit{R} & 12-Aug-16 \\
%			46 & 3751463 & 20-Sep-10 & spring & 3.0.5  & ScheduledTasksBeanDefinitionParser.java & 42 & 52 & 20-Oct-10 & \textit{S} & 21-May-12 \\
%			47 & 6149818 & 27-May-11 & spring & 3.0.5  & DefaultPropertiesPersister.java & 69 & 80 & 20-Oct-10 & \textit{S} & 19-Mar-13 \\
%			48 & 10952561 & 8-Jun-12 & spring & 3.0.5  & HibernateTemplate.java & 761 & 769 & 20-Oct-10 & \textit{S} & 9-May-15 \\
%			49 & 10952561 & 8-Jun-12 & spring & 3.0.5  & Jaxb2Marshaller.java & 253 & 269 & 20-Oct-10 & \textit{S} & 28-Aug-12 \\
%			50 & 20996373 & 8-Jan-14 & spring & 3.0.5  & test/../DelegatingServletInputStream.java & 6 & 20 & 20-Oct-10 & \textit{S} & 18-Dec-08 \\
%			51 & 20996373 & 8-Jan-14 & spring & 3.0.5  & web/../DelegatingServletInputStream.java & 6 & 20 & 20-Oct-10 & \textit{S} & 15-Jul-16 \\
%			52 & 20996373 & 8-Jan-14 & spring & 3.0.5  & servlet/../ DelegatingServletInputStream.java & 6 & 20 & 20-Oct-10 & \textit{S} & 15-Jul-16 \\
%			53 & 5660519 & 15-Apr-11 & spring & 3.0.5 & AnnotationMethodHandlerException\newline Resolver.java & 224 & 233 & 20-Oct-10 & \textit{D} & 20-Jan-12 \\
%			54 & 4781746 & 24-Jan-11 & spring & 3.0.5  & DispatcherServlet.java & 91 & 103 & 20-Oct-10 & \textit{S} & 8-Aug-11 \\
%			55 & 3758110 & 21-Sep-10 & spring & 3.0.5  & DefaultAnnotationHandlerMapping.java & 78 & 92 & 20-Oct-10 & \textit{D} & 20-Jan-12 \\
%			56 & 9003314 & 25-Jan-12 & spring & 3.0.5  & WebDataBinder.java & 95 & 108 & 20-Oct-10 & \textit{S} & 15-Aug-10 \\
%			57 & 14019840 & 24-Dec-12 & struts2 & 2.2.1  & DefaultActionMapper.java & 128 & 144 & 17-Jul-10 & \textit{S} & 18-Oct-13 \\
%			58 & 15131432 & 28-Feb-13 & struts2 & 2.2.1  & FreemarkerManager.java & 163 & 177 & 17-Jul-10 & \textit{S} & 28-Oct-13 \\
%			59 & 15110171 & 27-Feb-13 & struts2 & 2.2.1  & StringLengthFieldValidator.java & 25 & 42 & 17-Jul-10 & \textit{S} & 15-Jun-15 \\
%			60 & 21734562 & 12-Feb-14 & tomcat & 7.0.2  & BasicAuthenticator.java (2 overlaps) & 25 & 73 & 4-Aug-10 & \textit{R} & 4-Aug-16 \\
%			62 & 21734562 & 12-Feb-14 & tomcat & 7.0.2  & FormAuthenticator.java & 51 & 61 & 4-Aug-10 & \textit{R} & 4-Aug-16 \\
%			63 & 24404964 & 25-Jun-14 & tomcat & 7.0.2  & CoyoteAdapter.java & 543 & 553 & 4-Aug-10 & \textit{S} & 25-Sep-14 \\
%			64 & 10289462 & 23-Apr-12 & tomcat & 7.0.2  & JspRuntimeLibrary.java & 252 & 296 & 4-Aug-10 & \textit{S} & 12-Sep-12 \\
%			\hline
%	\end{tabular} %
%}
%\end{table*}

\begin{table*}
	\centering
	\caption{Examples of the outdated QS online clones}
	\label{tab:stale_code_details}
	\small
	\resizebox{2\columnwidth}{!}{%
		\begin{tabular}{l|l|l|l|p{4.7cm}|r|r|l|c|l}
		 \hline
		 \multicolumn{2}{c|}{Stack Overflow} & \multicolumn{6}{c|}{Qualitas} & \multicolumn{2}{c}{Changes} \\ 
		 \hline
		 Post & Date & Project & Ver. & File & Start & End & Date & Type$^*$ & Date \\
		 \hline
		 22315734 & 11-Mar-14 & hadoop & 1.0.0 & WritableComparator.java & 44 & 54 & 25-Aug-11 & \textit{S} & 20-Nov-14 \\
		 23520731 & 7-May-14 & hibernate & 4.2.2 & SchemaUpdate.java & 115 & 168 & 22-May-13 & \textit{S} & 5-Feb-16 \\
		 21734562 & 12-Feb-14 & tomcat & 7.0.2  & FormAuthenticator.java & 51 & 61 & 4-Aug-10 & \textit{R} & 4-Aug-16 \\
		 12593810 & 26-Sep-12 & poi & 3.6 & WorkbookFactory.java & 18 & 28 & 7-Dec-09 & \textit{R} & 26-Jun-13 \\
		 8037824 & 7-Nov-11 & jasperreports & 3.7.4 & JRVerifier.java & 1221 & 1240 & 31-May-10 & \textit{D} & 20-May-11 \\
		 3758110 & 21-Sep-10 & spring & 3.0.5  & DefaultAnnotationHandlerMapping.java & 78 & 92 & 20-Oct-10 & \textit{D} & 20-Jan-12 \\
		 \hline 
		 \multicolumn{10}{l}{{\footnotesize * \textit{S}=modified/added/deleted statements, \textit{D}=file deletion,  \textit{R}=method rewriting}}
		\end{tabular} %
	}
\end{table*}

\subsection{RQ4: Software Licensing Violation}
Software licensing plays an important role in software development. Violation of software licenses impacts software delivery and also leads to legal issues~\cite{Sprigman2015}. 
%It is an emerging area that software engineering research community is paying attention to. For example, 
One can run into a licensing issue if one integrates third-party source code into their software without checking. A study by An et al.~\cite{An2017} reports 1,219 cases of potential license violations between 399 Android apps and Stack Overflow code snippets. %there are studies of automatic technique to identify software licensing from source code files~\cite{German2010}

%\begin{table}
%	\centering
%	\caption{Licenses of online clones (file-level) from Qualitas projects}
%	\label{t:q_projects_license}
%	%	\resizebox{\columnwidth}{!}{%
%	\small
%	\begin{tabular}{l|l|r|r|r}
%		\hline 
%		Project & License & QS & UD & Total \\
%		\hline
%		netbeans &  CDDL/GPLv2 &  & 57 & 57 \\
%		&  No license &  & 8 & 8 \\
%		&  spdxBSD3 &  & 6 & 6 \\
%		&  MPLv1.1 &  & 1 & 1 \\
%		&  Unknown &  & 1 & 1 \\
%		\hline
%		jung2 &  BSD3 & 2 & 27 & 29 \\
%		\hline
%		eclipse-SDK & EPLv1 & 6 & 11 & 17 \\
%		& CDDL/GPLv2 & 1 & 1 & 2 \\
%		&  Apache-2 &  & 1 & 1 \\
%		& No license & 1 &  & 1 \\
%		& Unknown &  & 1 & 1 \\
%		\hline
%		hibernate & LesserGPLv2.1+ & 18 & 4 & 22 \\
%		\hline
%		spring & Apache-2 & 18 & 2 & 20 \\
%		\hline
%		vuze & GPLv2 &  & 11 & 11 \\
%		\hline
%		jstock & GPLv2+ &  & 9 & 9 \\
%		& Unknown &  & 1 & 1 \\
%		\hline
%		tomcat & Apache-2 & 5 & 5 & 10 \\
%		\hline
%		hadoop & Apache-2 & 8 &  & 8 \\
%		& No license & 1 &  & 1 \\
%		\hline
%		jfreechart & LesserGPLv2.1+ & 3 & 5 & 8 \\
%		\hline
%	\end{tabular} %
%	%	}
%\end{table}

\begin{table}[]
	\centering
	\caption{License mapping of the online clones (file-level) in QS, UD, and EX}
	\label{tab:license_abc}
	\small
	\resizebox{\columnwidth}{!}{%
	\begin{tabular}{l|p{2cm}|l|r|r|r}
		\hline
		Type & Stack Overflow\newline (CC BY-NC-SA) & Qualitas & QS & UD & EX \\
%		\hline
%		No license & No license & No license &  5 & 21 & 4 \\
		\hline
		Compatible & Apache2 & Apache2 &  0 & 1 & 0 \\
		License	& EPLv1 & EPLv1 &  1 & 1 & 0 \\
		& Sun (proprietary) & Sun (proprietary) & 0 & 0 & 1 \\
		& MIT-like & MIT-like & 0 & 0 & 1 \\
		& No license & CC BY-NC-SA &  0 & 0 & 5 \\
		& No license & No license &  5 & 21 & 4 \\
		\hline
		\multicolumn{3}{l|}{Total} &  6 & 23 & 11 \\
		\hline
		\hline
	    Incompat. & BSD3 & GPLv2+ & 0 & 1 & 0 \\
		License	& No license & AGPLv3/3+ & 1 & 2 & 0 \\
		& No license & Apache-2 & 41 & 23 & 5 \\
		& No license & BSD3 & 3 & 28 & 4 \\
		& No license & CDDL or GPLv2 & 1 & 58 & 0 \\
		& No license & CDDLv1 & 0 & 0 & 1 \\
		& No license & EPLv1 & 7 & 10 & 0 \\
		& No license & GPLv2/2+/3+ & 2 & 32 & 62 \\
		& No license & LGPLv2/2.1+/3+ & 24 & 32 & 0 \\
		& No license & MPLv1.1 & 0 & 1 & 0 \\
		& No license & spdxBSD3 & 0 & 10 & 0 \\
		& No license & SPLv1 & 0 & 0 & 1 \\
		& No license & Unknown & 0 & 3 & 3 \\
		& spdxBSD3 & AGPLv3+ & 0 & 2 & 0 \\
		\hline
		\multicolumn{3}{l|}{Total} &  79 & 202 & 76 \\
		\hline
	\end{tabular} %
}
\end{table}

%\begin{figure}
%	\centering
%	\includegraphics[width=\linewidth]{license_violation}
%	\caption{Summary of online clones' licensing conflicts}
%	\label{fig:license_violation}
%\end{figure}

%\begin{figure*}
%		\centering
%		\begin{minipage}{.4\textwidth}
%			\centering
%			\includegraphics[width=\linewidth]{sankey_license_qs}
%			\caption*{(A) QS clones}
%		\end{minipage}%
%		\hspace{5ex}
%		\begin{minipage}{0.4\textwidth}
%			\centering
%			\includegraphics[width=\linewidth]{sankey_license_ex}
%			\caption*{(B) EX clones}
%		\end{minipage}
%	\caption{Direction of the changes in online code clone licenses}
%	\label{fig:sankey_license}
%\end{figure*}

In our study, we reveal another possible situation of software licensing issues caused by code cloned ``to Stack Overflow''. We found evidence that 112 pieces of code had been copied from Qualitas projects to Stack Overflow as examples. Their status of accepted answers increase their chances of being reused. Even though most of the Qualitas projects came with a software license (shown in \Cref{t:q_projects_license}), we found that the licensing information were frequently missing after the code was copied to Stack Overflow. The licensing terms on top of source code files are not copied because usually only a small part of the file was cloned. %If developers copy and reuse these licensed pieces of code in their projects, conflicts may happen without their realisation.

We analysed licensing conflicts of the online clones in the QS, UD, and EX set. The summary of licensing information is listed in \Cref{tab:license_abc}. The licenses were extracted by \textbf{Ninka}, an automatic license identification tool~\cite{German2010}. Since Ninka works at file level, we report the findings based on Stack Overflow snippets and Qualitas source files instead of the clone pairs (duplicates were ignored). For the ones that could not be automatically identified by Ninka and have been reported as {\small\texttt{SeeFile}} or {\small\texttt{Unknown}}, we looked at them manually and updated if any license was found. The table shows the license mapping of Stack Overflow snippets and Qualitas source code. The results are reported separately for the QS, UD, and EX set. In overall, we can see that most of the Stack Overflow snippets do not contain licensing terms while their clone counterparts in Qualitas projects do.

\textbf{Compatible license:} There are 37 pairs which have compatible licenses such as Apache license v.2; Eclipse Public License v.1 (EPLv1); or a pair of No license vs.~Creative Common Attribution-NonCommercial-ShareAlike 3.0 Unported (CC BY-NC-SA 3.0). These clones are safe for being reused. Since source code and text on Stack Overflow posted before 1st March 2016 are protected by CC BY-NC-SA 3.0 (and MIT license from 1st March 2016 onwards), we can treat the Stack Overflow code snippets without licensing information as having CC BY-NC-SA 3.0 by default. CC BY-NC-SA 3.0 license is relaxed and it only requests an attribution when reused.

\textbf{Incompatible license:} there are 357 clone pairs which do not contain licensing information after they are posted on Stack Overflow or contain different license from their Qualitas clone counterparts. More than half (79) of QS clone pairs have their licensing terms removed or changed after posted on Stack Overflow. Of 226 UD clone pairs, 202 pairs have incompatible licenses. Again, most of the clones in Qualitas contain a license while the Stack Overflow snippets do not. For EX clone pairs, we searched for licensing terms of the original source code from the external sources. We found that 76 out of 87 EX clone pairs have incompatible licenses. %They are clones that have been identified to be copied from an external source. 
Similarly, the license statement was removed from Stack Overflow snippets. 
%
%\Cref{fig:sankey_license} shows the direction of changes in online code clone license. The direction of the change is from left to right. The thickness of the line reflects the number of clones for that relationship. In \Cref{fig:sankey_license} (A), we can clearly see that most of the QS online clones have their licenses changed from having a license to ``No license'' on Stack Overflow. The same findings were found for EX online clones as depicted in \Cref{fig:sankey_license} (B). We found that most of the EX clones have their licensing information stripped of after copying to Stack Overflow. These clones are instead covered by the default CC BY-NC-SA 3.0 license of Stack Overflow which is strongly permissive. These used-to-be-licensed code are dangerous for being reused because they can implicitly conflict with the reused software's license.

\textbf{For RQ4, we found 357 code snippets on Stack Overflow that violate the license of their original software. The majority of them do not contain licensing statements after copied to Stack Overflow.}

\section{Threats to Validity}

\textbf{Internal validity:} 
We could not verify all the 315,786,118 clone pairs reported by the clone detection tools in this study.
Nevertheless, we applied different mechanisms to ensure the validity to the subset of clone pairs we classified.
First, we used two, instead of one, widely-used clone detection tools, Simian and NiCad. 
We had tried three other clone detectors but could not add them to the study
due to their scalability issue and susceptibility to incomplete code snippets.
Second, we adopted Bellon's clone agreement metric~\cite{Bellon2007} to prioritise and filters clone pairs 
for the manual classification. The agreed clone pairs that passed \textit{good} and \textit{ok} criteria had a higher chance
to be true positive. We also applied the Bellon's clone agreement to 4 different combinations of the tools' configurations.
For disagreed clone pairs, we applied several filters to remove spurious and uninteresting clones.

The number of online clones we found are subject to the clone detectors and their configurations. 
We mitigated the problem of clone detector's parameter sensitivity by adding another configuration, besides the default one. We reused the established configurations from the EvaClone study~\cite{Wang2013}.
We also performed a pairwise matching between the default and EvaClone configurations of the two clone detection tools.

Our seven patterns of online code cloning may not cover all possible online cloning patterns. However, instead of 
defining the patterns beforehand, we resorted to extract them from the data sets. We derived them
from a manual investigation of 679 online clone pairs and adopted one pattern from the study by Kapser et al.~\cite{Kapser2003}.

The 3,632 clone pairs classified manually by the first author are subject to manual judgement and human errors. 
Although we have tried our best to be careful on searching for evidence and classifying the clones, some errors
may still exist. We mitigated this problem by having the second author to validate the classifications with random samplings.
The 64 classification conflicts were discussed and resolved. 
The third author made a final judgement when no consensus could be found by the two authors.
 %In addition, we are aware of voting mechanism
%in Stack Overflow and plan to incorporate highest-vote answers in the future work.

\textbf{External validity:} We carefully chose the data sets for our
experiment so the findings could be generalise as much as possible. 
We selected Stack Overflow because it is one of the most popular programming Q\&A websites available with approximately 6.3 million users.
There are a large number of code snippets reused from the site~\cite{An2017} and there are also several studies encouraging of doing so~(e.g.~\cite{Ponzanelli2013,Ponzanelli2014,Keivanloo2014,Park2014}). 
Nonetheless, it may not be representative to all the programming Q\&A websites.

Regarding the code snippets, we downloaded a full data dump and extracted Java accepted answers. Instead of analysing every post, we only focused on posts marked as accepted answers 
since they are the most likely to being reused. This yields a total of 144,064 snippets.
Our findings are limited to these restrictions. They may not be generalised to all programming languages and all answers on Stack Overflow. On the other hand, we chose the curated Qualitas corpus for Java open source projects containing 111 projects \cite{QualitasCorpus}.
The projects span over several areas of software and has been used in several empirical studies~\cite{Taube-Schock2011,Beckman2011,Vasilescu2011,Omar2012}. Although it is a curated and well-established corpus, 
it may not fully represent all Java open source software available.

%Lastly, two clone detection tools, Simian and NiCad, are chosen for this study. We have not tried all the tools available and they might not be complete representatives of all available clone detection tools.

\section{Related Work}

\textbf{Stack Overflow} is a gold mine for software engineering research. Its rich and developer-driven data are invaluable. Since posts on Stack Overflow may contain code snippets embedded within natural language text, they become a huge database for source code and code-relevant information. The Stack Overflow data set has been put to use in several previous studies. In terms of developer-assisting tools, Seahawk is an Eclipse plug-in that searches and recommends relevant code snippets from Stack Overflow~\cite{Ponzanelli2013}. A follow up work, Prompter, by Ponzanelli et al.~\cite{Ponzanelli2014} achieves the same goal but with improved algorithms. The code snippets on Stack Overflow are mostly examples or solutions to programming problems. Hence, several code search systems use whole or partial data from Stack Overflow as their code search databases~\cite{Keivanloo2014,Park2014, Stolee2014,Subramanian2013,Diamantopoulos2015}. Furthermore, Treude et al.~\cite{Treude2016}~use machine learning techniques to extract insight sentences from Stack Overflow and use them to improve API documentation.

Another research area is knowledge extraction from Stack Overflow. Nasehi et al.~\cite{Nasehi2012}~studied what makes a good code example by analysing answers from Stack Overflow. Similarly, Yang et al.~\cite{Yang2016} report the number of reusable code snippets on Stack Overflow across various programming languages. Wang et al.~\cite{Wang2013_StackOverflow} use Latent Dirichlet Allocation (LDA) topic modelling to analyse questions and answers from Stack Overflow so that they can automatically categorise new questions. There are also studies trying to understand developers' behaviours on Stack Overflow, e.g.~\cite{Movshovitz-Attias2013,Rosen2016,Choetkiertikul2015,Bosu2013}, while some studies aim to improve Stack Overflow itself~\cite{Diamantopoulos2015, Wang2014}.

\textbf{Code clone detection} is a long-standing research topic in software engineering. Whether clones are good or bad for software is still controversial~\cite{Sajnani2016,Kapser2003,Kapser2008,Krinke2008,Hotta2010,Gode2011,Harder2013}. However, by only knowing how many code clones residing in software and how they evolve~\cite{Pate2013,Mondal2011} can provide several valuable insights into the software systems. Code clones have several applications such as software plagiarism detection~\cite{Prechelt2002}, 
source code provenance~\cite{Davies2013}, and software licensing conflicts~\cite{German2009}.

Two code fragments are clones if they are similar enough according to a given definition of similarity~\cite{Bellon2007}. Given an open interpretation of ``definition of similarity'', there are various clone detection tools and their siblings, code plagiarism detectors, invented based on plethora of different code similarity measurements~\cite{Roy2008, Ragkhitwetsagul2016,Svajlenko2014}. Some tools
use string comparison techniques such as Simian~\cite{simian}. NiCad~\cite{Roy2008,Cordy} exploits the Longest Common Subsequence (LCS) string similarity measure after applying code pretty-printing using TXL~\cite{Cordy2006}. Many tools do not work on original source code directly but transform them into an
intermediate representation such as tokens and apply similarity
measurement on them. These tools include SourcererCC~\cite{Sajnani2016}, CCFinder~\cite{Kamiya2002},
CP-Miner~\cite{Li2006}, iClones~\cite{Gode2009} and a few more~\cite{Burrows2007, Smith2009, Duric2012, Prechelt2002, Schleimer2003}. 
To find more challenging clones such as clones with added/deleted/reordered statements or equivalent loop and conditional statements (i.e.~type-3 clones), structural similarity of clones is needed.
This structural similarity can be discovered by comparing AST as found in CloneDR~\cite{Baxter1998} and Deckard~\cite{Jiang2007a} or by using program dependence
graphs~\cite{Krinke2001,Komondoor2001}.

Kapser et al.~studied clones in Linux file systems and create 11 patterns of code cloning based on four groups: Forking, Templating, Customisation and Exact match~\cite{Kapser2003,Kapser2008}. Our study adopted one of the patterns into our online code cloning patterns.

\textbf{Clone agreement} is useful when a clone oracle is absent. Since clone detectors are different in their detection approaches, they may behave differently and report different clones even on the same data set. Some researchers exploit these different behaviours of clone detectors by finding their agreement and obtain highly-confident clones~\cite{Bellon2007,Wang2013}. Using the same data set, clone pairs that are agreed by several tools (or several code similarity measurement techniques) are more assured to be true clones than the ones reported by only a single tool~\cite{Wang2013,cr2016ssbse,Funaro2010}. There are studies that report sensitivities of the tools' configurations to the results~\cite{Wang2013,cr2016ssbse}, showing that searching for an optimal configuration for every new data set is preferred.

\textbf{Software licensing} is crucial for open source and industrial software development. Di Penta et al.~\cite{DiPenta2010}~studied the evolution of software licensing in six FOSS and found that licensing statements change over time. German et al.~\cite{German2009} performed an empirical study of code siblings (code clones among different systems coming from the same source) and found that licensing conflicts can occur between the clone siblings. Later, German et al.~\cite{German2010} created Ninka, a tool to automate the identification of software license from program source code. We use Ninka to analyse software license from Stack Overflow snippets and Qualitas projects.

\textbf{Reusing of outdated third-party source code} occurs in software development. Xia et al.~\cite{Xia2014} show that a large number of open source systems reuse outdated third-party libraries from popular open source projects. Using the outdated code give detrimental effects to the software since they may introduce vulnerabilities. Our study discovers similar findings in the context of outdated code on Stack Overflow.

The work that is closely similar to us is a study by An et al.~\cite{An2017} where the authors investigated clones between 399 Android apps and Stack Overflow posts. They found 1,226 code snippets which were reused from 68 Android apps. They also observed that there are 1,219 cases of potential license violations. The authors rely on the timestamp to judge whether the code has been copied from/to Stack Overflow along with confirmations from six developers. Instead of Android apps, we investigated clones between Stack Overflow and 111 open source projects. Their results are similar to our findings that there are clones from software projects to Stack Overflow with potential licensing violations. %In our work, we defined seven patterns of online code cloning and performed a large-scale manual check of 3,632 clone pairs. We discovered 112 clone pairs with strong evidences, based on natural text in comments and post contents, that they were copied from Qualitas to Stack Overflow. By comparing the clones to their latest versions in the software, we found that 64 code snippets on Stack Overflow are outdated and possibly harmful for reuse.

%\section{Future Work}
%Since this study is limited to only Java language, we are interested in expanding the study further to other popular programming languages such as C, C++, Python, and Javascript on Stack Overflow. The similar study of outdated code and software licensing violation can be repeated. It is interesting to compare the findings among different programming languages. Moreover, other popular software project corpora (e.g.~GitHub) can also be included to increase generalisation of the findings.
%
%We are eager to tackle the issues of outdated and license-violating online code clones with an automated tool. The tool should incorporate a scalable clone detection method of locate similar code from programming Q\&A websites such as Stack Overflow and online code repositories like GitHub. It analyses a software project under development and reports clones that are copied from Stack Overflow. The reported clones have traces to their origins in open source projects or websites along with their freshness (fresh/outdated/dead) so that the developer of the project can be aware of changes made to the copied code snippets. Lastly, the license of the original source code is also reported so the developer can make an informed decision of whether he or she will incorporate that code snippet into their own project.

\section{Conclusions}
Online code clones are clones that have been copied to Q\&A websites such as Stack Overflow. They are similar to clones in software projects where they can be outdated and violating software license but are much harder to detect. We studied the online code clones between Stack Overflow and Qualitas projects, especially 3,632 clone pairs, manually in details.

Although we deployed Bellon's clone agreement and multiple filters to reduce the number of clone candidates, we still found a large number of 51,926 online clone pairs reported by two clone detectors. We classify 31,946 clone pairs using our seven patterns of online code cloning. We discovered 112 clone pairs that have been copied, with evidence, from Qualitas projects to Stack Overflow (QS), 197 clone pairs that have been copied from external sources besides Qualitas to Stack Overflow (EX), and 359 clone pairs that are highly similar but without evidence of copying (UD).

A manual investigation of the 112 QS clone pairs by comparing to their latest version (as of 26 September 2016) reveals that 64 of them are outdated. Moreover, the study of licensing conflicts of clones in QS, EX, and UD shows that 357 code snippets on Stack Overflow do not contain licensing statements while their clone counterparts in Qualitas projects or external sources do.

This study is among, if not the first, to address the important issues of outdated and license-violating online code clones on programming Q\&A websites using a hybrid methodology of automatic code clone detection and a manual clone investigation.

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}

\end{document}