### 2018-05-24 cloverflow-tse/cover_letter/cover_letter.tex
After the Grammarly check.

% Copyright Javier Sánchez-Monedero.
% Please report bugs and suggestions to (jsanchezm at uco.es)
%
% This document is released under a Creative Commons Licence 
% CC-BY-SA (http://creativecommons.org/licenses/by-sa/3.0/) 
%
% BASIC INSTRUCTIONS: 
% 1. Load and set up proper language packages
% 2. Complete the paper data commands
% 3. Use commands \rcomment and \newtext as shown in the example

\documentclass[a4paper,twoside,10pt]{reviewresponse}

% 1. Load and set up proper language packages
%\usepackage[utf8x]{inputenc}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage[latin9]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{natbib}
\usepackage{float}
\usepackage{multirow}
\usepackage{booktabs} % For formal tables

% 2. Complete the paper data
\newcommand{\myAuthors}{Chaiyong Ragkhitwetsagul,~Jens Krinke,~Matheus Paixao,~Giuseppe Bianco,~Rocco Oliveto}
\newcommand{\myAuthorsShort}{Ragkhitwetsagul et. al}
\newcommand{\myEmail}{{chaiyong.ragkhitwetsagul.14,j.krinke,matheus.paixao.14}@ucl.ac.uk, giuseppebianco92@gmail.com, rocco.oliveto@unimol.it}
%\newcommand{\mySecEmail}{}
\newcommand{\myTitle}{Cover Letter for the Reviewers of the Paper \\ ``Toxic Code Snippets on Stack Overflow''}
\newcommand{\myShortTitle}{Cover Letter}
\newcommand{\myJournal}{IEEE TRANSACTIONS ON SOFTWARE ENGINEERING}
\newcommand{\myDept}{University College London (UK), University of Molise (Italy)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage[linktoc=all]{hyperref}
\usepackage[linktoc=all,bookmarks,bookmarksopen=true,bookmarksnumbered=true]{hyperref}

\hypersetup{
pdfauthor = {\myAuthorsShort},
pdftitle = {\myTitle},
pdfsubject = {\myJournal\xspace},
colorlinks = true,
linkcolor=blue,          % color of internal links
citecolor=black!70!green,        % color of links to bibliography
filecolor=magenta,      % color of file links
urlcolor=black           % color of external links
}

\newcommand\FIXME[1]{{\color{red}\textbf{FIXME: #1}}}

\begin{document}

\thispagestyle{plain}

\begin{center}
 {\LARGE\myTitle} \vspace{0.3cm} \\
 {\large\myJournal} \vspace{0.3cm} \\
 \today \vspace{0.3cm} \\
 \myAuthors \\
 \url{\myEmail} \\
 %\url{\mySecEmail} 
 \vspace{0.3cm} 
 \myDept \vspace{1cm}
\end{center}

%\tableofcontents

%\begin{abstract}
This paper is a journal-first submission to Transactions on Software
Engineering. This is a re-submission of the previous submission
no.~TSE-2017-11-0335 and we would like to thank the anonymous reviewers for
taking their time to read our paper carefully. With their useful and
constructive reviews, we substantially improved the work by adding five major
improvements as listed below to our previous submission and minor improvements
as discussed later in this letter.

\textbf{1.~A Stack Overflow Visitor survey}~To understand the effects of toxic
code snippets to information seekers who look for coding
solutions on Stack Overflow, we performed another survey of 87 Stack Overflow
visitors. The survey shows that the visitors experienced several problems from
reusing Stack Overflow code snippets including outdated code. They are 85\% of
them that are not aware of CC BY-SA 3.0 license enforced by Stack Overflow and
66\% never check for license conflicts when reusing code snippets.

\textbf{2.~A study of toxic code snippets on GitHub:}~To provide evidence that
toxic code snippets are actually reused in software projects and to strengthen
the motivation, we performed a large-scale analysis of clones between the
discovered online code clones (100 outdated and 214 license-incompatible code
snippets) and 130,703 GitHub projects using a clone detection tool followed by
manual validation.

\textbf{3.~An analysis of the tools' parameter settings:}~We mitigated the
threats to validity due to parameter settings of the clone detection tools and
the clone merging technique by exploring different tools' parameter values. We
discovered that using the minimum clone size threshold of 10 lines gave the
highest number of non-trivial clone pairs and using the Bellon's ok-match
threshold of 0.5 gave the highest clone merging precision. We updated the
results in the paper to reflect the findings.

\textbf{4.~An extended analysis of outdated code clones:}~We performed an
extended analysis of the 100 outdated code snippets. For example, we analysed
the age of the clones; studied effects of comments, newer answers, higher-voted
answers on the outdated answers; and summarised six intents that made the
clones outdated. We found that, on average, the outdated code snippets were cloned to Stack Overflow
2 years after it appeared in Qualitas projects, and comments/newer or
higher-voted answers did not strongly affect the outdated answers (99\% are
still marked as accepted). Lastly, we discovered that the code snippets are outdated
because of code enhancement (65\%), followed by code deprecation (15\%), but
fixing (10\%), code refactoring (6\%), changing coding style (3\%), and data
format change (2\%).

\textbf{5.~Revisions to the overall discussion:}~We substantially expanded the
overall discussion to include (1) a discussion of toxicity of outdated and
license-incompatible clones on Stack Overflow and their implications, and (2)
more detailed discussion of how to implement the tools or approaches proposed
for preventive and detective measures, and (3) a suggestion for automation of
code cloning direction detection.

\clearpage
We also tried our best to address all the comments from the reviewers. Every
comment led to an improvement on this submission, and we discuss them in order as
shown in the next sections.

Lastly, we would also like to ask the reviewers to comment on how to address the
newly established page limits of IEEE TSE that was enforced after our previous
submission. After incorporating the suggestions from the reviewers and
rewriting, our paper is now extended to 22 pages. Are there any elements that
could or should be removed, or should a split into two separate papers be
considered?

\clearpage

\section{Reviewer 1}

\rcomment{ Regarding this statement: \textit{``It is probably more important
		than studying the effects of reusing Stack Overflow code snippets because it
		gives insights to the root cause of the problem and lays a foundation to an
		automatic technique to detect and report toxic code snippets on Stack Overflow
		to developers in the future.''}

	While~\cite{An2017} do suggest that copying from Stack Overflow to software
	projects occurs (and results in license violations), they do not establish that
	code snippets from Stack Overflow \textit{``go viral''} and end up distributed
	widely across (open or closed source) software projects. Moreover, while the
	quote on the first page implies that developers often \textit{``mindlessly''}
	reuse code snippets from Stack Overflow, to the best of my knowledge, this claim
	is conventional wisdom but not empirically validated. So, it seems to me that
	the primary interest in toxic code snippets is that may cause widespread harm,
	but I am missing the evidence that they do so in practice. Even assuming that
	code snippets from Stack Overflow are distributed mindlessly and widely, I
	wonder how many code snippets on Stack Overflow are toxic, the extent to which
	they are reused without modification (or with only minor adaptation), what kinds
	of projects they end up in (i.e., a popular project from Microsoft or Apache vs.
	a hobby project by a single developer vs. a student solution to an assignment),
	and ultimately, how much damage can we expect them to cause. In the end, I agree
	that it is better to detect and report toxic code snippets than to not do so,
	but I am not sure how big of a problem is solved by doing so. As such, I
	disagree that it is \textit{``probably more important''} to find the toxic code
	snippets than to understand their effects. Instead, I think that understanding
	their effects (on software projects rather than on the quality of the Stack
	Overflow platform) is necessary to appreciate the value in finding them. }

We would like to thank the reviewer for the comments.~We agree with the
suggestion, and we made the following changes to the paper:

\begin{enumerate} \item We understand the concern of finding real evidence to
	support the claim that code snippets on Stack Overflow do appear in open- or
	closed-source software projects. Hence, we have performed a study of code clone
	detection between our 100 reported outdated code snippets on Stack Overflow and
	131,703 GitHub projects. We selected SourcererCC with 80\% similarity for this
	experiment because it could scale to a large-scale data set, while Simian could
	not. We found 102 cloned candidates appearing in 68 GitHub projects and manually
	investigated all of them. As shown in Table~\ref{tab:outdated_github} below, out
	of the 102 cloned snippets, 13 cloned snippets matched with
	themselves because some of the Qualitas projects also appear on GitHub. For
	other projects besides the Qualitas projects, 47 cloned snippets
	were exactly the same as the outdated code snippet and 42 cloned snippets that
	were not the same (e.g.,~they contained additional code modifications
	made by the projects' developers, or they were copied from another source with a
	slightly different code).

	\begin{table}[H]
		\centering
		\begin{tabular}{lr}
			\toprule
			Clones & Amount \\
			\midrule
			\textit{Found in Qualitas GitHub repos} & 13 \\
			\midrule
			\textit{Found in other project repos} & \\
			Exact copy (outdated) & 47 \\
			Non-exact copy & 32 \\
			\midrule
			Total & 102 \\
			\bottomrule
		\end{tabular}
		\label{tab:outdated_github}
		\caption{Clones of the 100 Stack Overflow outdated code snippets in 131,703 GitHub projects}
	\end{table}

Among the exact-copy of 47 outdated cloned snippets, there were only two cloned
snippets that gave attributions to Stack Overflow. However, the attributions
pointed to different posts than the ones we found but containing the same code
in the answers. The other 32 cloned snippets were very likely to be a
file-level clone from its respective original project (e.g.,~JFreeChart, JUnit,
Log4J, Hadoop) based on their license header and the Javadoc comments. Lastly,
the remaining 13 cloned snippets did not have any hints or evidence of copying.
Twelve of them were toxic because they were the old versions that still
contained bugs.

Interestingly, we discovered that the buggy version of the
\texttt{humanReadableInt()} method from Hadoop appears in two popular Java
projects: deeplearning4j (8,830 stars and 4,223 forks) and Apache Hive (1,854
stars and 1,946 forks). Due to the lack of evidence, we could not conclude how
this method, which is exactly the same as the toxic code snippet we found on
Stack Overflow, appear in the two projects. It is possible that the developers
retrieved them from Stack Overflow, other websites, or from Hadoop code base
directly. Nevertheless, we reported them to the developers of the two projects
regarding the issue. We created a bug report for each project (deeplearning4j
\#4694\footnote{deeplearning4j bug report:
	\url{https://github.com/deeplearning4j/deeplearning4j/issues/4692}} and
HIVE-18929\footnote{Apache Hive bug report: \url{
		https://issues.apache.org/jira/browse/HIVE-18929}}) and communicated with the
developers by describing the problem of race condition in the outdated version
of the \texttt{humanReadableInt()} method and proposed a fix by using the
newest version of the method in Hadoop. The issue has been fixed in
deeplearning4j. The developers of deeplearning4j agreed that the method was
problematic and decided to implement their own fix due to a concern of a
potential software licensing conflict caused by copying the fix directly from
the Hadoop code base. On the other hand, the bug report in Apache Hive is still
open.

	\item The same study has been done for another type of toxic code snippets on
Stack Overflow, license-incompatible code snippets. We detected clones between
the 214 code snippets without their original license (86 QS, 78 EX, and 50 UD) 
and 130,703 GitHub projects using SourcererCC with 80\% similarity threshold.
In contrast to the outdated clones, we discovered many more clones of 7,207
candidate pairs. 
There were 95 pairs from 10
Qualitas projects hosted on GitHub and 7,112 pairs from 2,427 other projects.
Then, we used the Ninka tool to identify
software licenses from the reported cloned snippets automatically. Shown in
Table \ref{tab:license_github} below, the clones were found in highly-starred
projects (29,465 to 10 stars) to 1-star projects. We found 12 code snippets
with attributions to Stack Overflow questions/answers and 6 of them were the
answers we found as clones from Qualitas or external sources. Five code
snippets did not have a license while the originals had the Apache-2, GPLv2, or EPLv1
license. One snippet had the AGPLv3 license while the original had the Apache-2
license. Only 995 code snippets in GitHub projects have the same license
as the originals in Qualitas.

	\begin{table}[H]
		\centering
		\begin{tabular}{l|rr|rrr}
			\toprule
			\multirow{2}{*}{No. of stars} & \multicolumn{2}{c|}{Qualitas} & \multicolumn{3}{c}{Other Projects} \\ \cmidrule{2-6}
			& Projects & Clone Pairs & Projects & Clone Pairs & Same license \\
			\midrule
			29,540 to 10 stars & 8 & 71 & 406 & 1,837 & 193 \\
			9 to 5 stars & 0 & 0 & 275 & 739 & 110 \\
			4 to 1 stars & 2 & 24 & 1,746 & 4,536 & 692 \\
			\midrule
			Total & 10 & 95 & 2,427 & 7,112 & 995 \\
			\bottomrule
		\end{tabular}
		\label{tab:license_github}
		\caption{Clones of the 214 Stack Overflow missing-license code snippets in 131,703 GitHub projects}
	\end{table}

	\item With the finding above, we show that clones of toxic Stack Overflow 
code fragments do occur in several open source software projects on GitHub.
However, we could not establish strong evidence that the toxic code snippets
were originated from Stack Overflow due to lack of supporting information.
Hence, we softened our claim by changing the statement ``\textit{It is probably
	more important than studying the effects of reusing Stack Overflow code
	snippets ...}'' to ```\textit{It is \textbf{equally} important to studying the
	effects of reusing Stack Overflow code snippets ...}''. We also updated the
Overall Discussion section to reflect the findings that (1) there is only a
small number of toxic code snippets from outdated code that are found in GitHub
projects, (2) there is a large number of toxic code snippets caused by a missing
license that are found in GitHub projects. %Moreover, the direction of copying
% is also unclear. The clones could come from Stack Overflow, the original
% project repositories, or other places.
\end{enumerate}

\rcomment{
	I think that the survey evidence is compelling. But I am missing the awareness
of Stack Overflow questioners or information-seekers to toxic code snippets. I
wonder: how skeptical are developers when considering code snippets on Stack
Overflow, how often do developers use code snippets directly, and how often do
developers use clues such as \textit{``A quick google search returned me this
	from Appache hadoop project. Copying from there:''} to seek out the latest
version of the information provided in the answer. The survey responses
described in Section 3.5.3 suggest that there are
questioners/information-seekers that are aware of toxic code snippets (as they
are reporting them to the answerers), and there is evidence that the Stack
Overflow community is aware of the need to retire outdated information and of
the many concerns related to software licensing, as these issues are discussed
on meta.stackoverflow.com. So, again, I wonder about the impact of toxic code
snippets on actual software projects.
}

We partially address this comment in the previous answer by performing an
empirical study of code cloning between toxic Stack Overflow code snippets and
GitHub projects. Furthermore, we also improved the paper on the aspect of the
awareness of information-seekers to toxic code snippets on Stack Overflow by
including the results from another survey of Stack Overflow visitors
(i.e.,~information seekers). This survey of the Stack Overflow visitors was
performed during the same time as the Stack Overflow answerers (25 July 2017 to
25 September 2017), and the full results are reported in our technical
report~\citep{Ragkhitwetsagul_RN2017}. We did not include the results in the
previous version of the paper due to a smaller sample size compared to the
answerers' survey. However, we did find some interesting results, so we have
included them in this version and discussed the key findings in~RQ1. The brief
summary of the visitor survey is as follows.

\textbf{Stack Overflow visitors' survey}

We surveyed 87 Stack Overflow visitors (visited the site and reused code from Stack
Overflow at least once) via five channels: social media post (Facebook),
\textsf{blognone.com} (a popular IT community and discussion forum in Thailand),
the University of Molise in Italy where the third author works, the
\texttt{comp.lang.java.programmer} group and the Software Engineering Facebook
page.

The visitors' survey confirms the findings from the previous studies that Stack
Overflow code snippets can be problematic~\citep{Zhang2018,Acar2016,An2017}. Sixty-six
percent of the visitors experienced a problem from reusing Stack Overflow code
snippets ranging from incorrect solutions, outdated solutions, mismatched
solutions, to buggy code. Although they are aware of the problems, half of them
(56\%) never report back to the Stack Overflow discussions. On the other hand,
the visitors rarely give attributions to Stack Overflow when they reuse code
snippets from the website, similar to the findings reported
by~\cite{Baltes2017}. The visitors are generally not aware of the CC BY-SA 3.0
license, and more than half of them never check for license compatibility when
reusing Stack Overflow code snippets. 
Sixty-nine Stack Overflow visitors (79\%) who adopted code from Stack Overflow never 
check if the code snippet originated from a different source (e.g.,~an open source project) 
and if they had an incompatible license to their projects.
Moreover, we also found that 9\% of the participants
encountered legal issues by copying code from Stack Overflow. To the best of our
knowledge, a study of legal problems from reusing Stack Overflow code snippets
has never been done before and will be our future work.

The full detail can be found in the RQ1 result section in the paper.

\rcomment{ The \textbf{Overall Discussion} section is not commensurate with the
	quality and breadth of the empirical study. The proposed preventative measure is
	actually two measures: a policy change and an IDE plugin. While an IDE plugin
	would provide early detection (and possible prevention) of license violations,
	the exact mechanics are not clear based on the description. Further, companies
	(and perhaps major open source projects?) use auditing platforms/services to
	detect such violations. Regarding the detective measure, I do not think that the
	paper establishes that a system to detect outdated code snippets on Stack
	Overflow is needed (but rather it establishes that outdated code snippets exist
	on Stack Overflow). Further, not all outdated code snippets are bad. For
	instance, a code snippet may show a solution using an older version of an API,
	and that older version may still be used in production. So, the definition of
	outdated would need to be a bit more nuanced to be used in an automated system
	for flagging or updating old Stack Overflow posts. }

Thanks for the suggestion.~We rewrote the Overall Discussion section to reflect
yours and Reviewer 3's comments. The changes include:

\begin{enumerate} \item We added a discussion of toxicity of outdated and
	license-incompatible clones on Stack Overflow. Although we found some harmful
	outdated clones, we could not find evidence that they are widely reused in open
	source projects on GitHub. Therefore, the damage is not high since there are only
	a few outdated code snippets that are harmful to reuse. On the other hand, the
	problem of license incompatibility caused by missing original licenses in Stack
	Overflow code snippets can cause more damage than outdated code. Different legal
	system and different interpretation of fair-use concept and de minimis (i.e.,~the
	minimal lines of code allowed to be copied without violating copyright) in each
	country can make reuse of these code snippets legally problematic.

	\item We improved the Actionable Items to be more precise about the mechanics
	on how to automate the process of finding online code clones, detecting
	outdated code, and finding license-incompatible code snippets.

	\item To address your point about auditing platforms/services, we performed a
	study of two open source software auditing platforms/services:~BlackDuck
	Software\footnote{https://www.blackducksoftware.com} and
	nexB\footnote{https://www.nexb.com}. For BlackDuck Software, we found from
	their report~\cite{CORSI2017} that while they check for code copied from open
	source projects including GitHub and Stack Overflow and analyse their license
	compatibility with their customer software, the BlackDuck auditing system will
	treat the code snippets on Stack Overflow as having an ``unknown'' license because
	it does not know the original license of Stack Overflow code snippets. For
	nexB, their product does not mention checking of reused source code from Stack
	Overflow. So, our proposed preventive measure, which can offer more precise
	licensing information of Stack Overflow code snippets, will be useful as an
	add-on license check for code copying from Stack Overflow. We include this in
	the Overall Discussion. \end{enumerate}

\rcomment{ Other Comments While ``Toxic Code Snippets'' is a catchy phrase for
	the title, I think that online code clones are the dominant topic.}

You are right. We mainly study online code clones and established their
existence. However, we used the term ``toxic code snippets'' to
emphasise the possible ramifications from online code cloning. Moreover, the
term code cloning is not well known outside the code cloning community. Code
clones are called differently in each research community or practice, such as
duplicated code, copy-and-paste code, code reuse, software similarity or
software redundancy. So, we decided not to use the term in the title.

We also updated the Introduction to state that the dominant topic is about
online code clones as follows.

``Although this activity of online code cloning is well-known, there are only a
few empirical studies on the topic~\citep{An2017,Abdalkareem2017,Baltes2017},
especially on finding the origins of the clones. We tackle this challenge of
establishing the existence of online code clones and investigate how they occur in
this study.'' %Considering your suggestion and another suggestion by Reviewer 2
% that the definition of ``toxic code snippets'' is vague, we also tightened the
% definition of toxic code snippets to be more precise.

\rcomment{ The introduction states that the erroneous code snippet in post
	22262310 has been viewed 259 as of the writing of the paper. But I note that
	while the question was posted on Mar 7, 14 the answer was posted on Mar 11, 14.
	I also note that the question has a score of -1. So, I wonder (a) whether the
	view count is for the answer (i.e., the code snippet) or the post (which existed
	for a few days before the answer appeared), (b) how many people viewed the
	question before it was answered, and (c) if the negative score for the question
	impacts both the number of views and the trust that people place in the answer
	to a down-voted question.}

That is a fair point, and we would like to analyse the statistics as you
suggested. To answer your question (a), the number of views is for the whole
Stack Overflow post but we use it as a proxy of the number of views the accepted
answer receives because the question and the answer of the two motivation
examples have a short gap of posting time (within the same day and four days
after). For (b), unfortunately, for our experiment we extracted only the code
snippets from the Stack Overflow dumps but not other information. They are the
only available data of the January 2016 snapshot we currently have. We checked
and found that the latest available Stack Overflow dump already contains
different data than the one we used in our study since it is more than 2
years newer, and will result in incorrect statistics and inconsistent results.
For (c), that is a really interesting idea. However, it is out of the scope of
this study and would involve extensive analysis of the data so we may
investigate it in the future work.

Nevertheless, we have added a footnote to the text in the Introduction to
clarify the point (a).

\rcomment{ FYI, since this paper was authored, a Stack Overflow user has posted
	a comment in response to the answer for post 22262310. The comment highlights
	the missing line and refers to the relevant HADOOP issue report. Interestingly,
	the gap between the issue report being resolved (Nov 20, 14) and the comment
	appearing on Stack Overflow (Oct 31, 17) is nearly three years. I wonder how
	many views occurred during this gap and whether the addition of the comment
	mitigates the toxicity of the code snippet. }

Thank you for the detailed checking of the Stack Overflow post. Interestingly,
this is surprising to us as well. The comment in the post 22262310 by Joseph K
on 31 October 2017 that highlights the missing line in HADOOP \texttt{compare}
method is actually from a student in our class. His name is Joseph Kalash, and he
is a master student in Machine Learning MsC program at UCL Computer Science
department\footnote{Joseph's CV can be found here: \url{https://jkala.sh/cv}}.
We assigned a preprint of this paper to students in the COMPGS11 -- Research
Seminar in Software Engineering course and had a discussion about the paper on
Thu 2 November 2017. Joseph, who was an active Stack Overflow member, was one of
the students in the class and he actively participated in the discussion. It
seems that he had seen the Hadoop's toxic code snippet in the paper and decided
to post a comment to warn other users.

Nevertheless, since this is a very interesting suggestion, we also performed an
investigation of the 100 answers of our reported outdated code snippets on Stack
Overflow on 6 May 2018 to see if there is any comment to mitigate or point out
the toxicity of the code snippet. We include the result of this investigation in
RQ4. We found that out of 100 answers containing outdated code snippets,
6 answers had a comment saying the code in the answer is
outdated or contains issues (including Joseph's comment). Thus, there are about
6\% of outdated answers that were notified by the Stack Overflow users based on
our analysed data. None of the 6 answers change their outdated code, but the
comments themselves may help to warn other Stack Overflow users. We added this
discussion to RQ4.

\rcomment{ \begin{enumerate} 
		\item Table 1 has a median column. Median of what?
		\item Section 2.1 states that comments were removed before clone detection, but
		as stated later in the paper, comments can provide clues regarding provenance.
		Can the manual inspection of comments to find evidence of copying be automated
		as part of the clone detection process? 
		\item Typo: ``how much the community
		trust them'' --> ``how much the community trusts them'' \end{enumerate} }

\begin{enumerate} 
	\item For the Stack Overflow data set, the value represents
	the median size of Stack Overflow code snippets. For the Qualitas data set, the
	value represents the median of the number of lines of code in each project. Since
	they are confusing as you pointed out, we decided to remove them from the table
	and only discuss the numbers in the text. 
	\item This is a great insight! We expanded the idea further and came up with a suggestion for an
	automated technique to facilitate human investigator on finding provenance and
	included it in a section called Automated cloning direction detection in the
	Overall Discussion. 
	\item Thanks. This has been fixed. \end{enumerate}

\section{Reviewer 2}

\rcomment{ \textbf{The definition of toxic code snippets:} I understand that the
	authors are trying to put a new spin or name to the phenomena they study, but
	the definition the authors provide is, in my opinion, not sufficient. For
	example, their definition is ``toxic code snippets mean code snippets that are
	harmful to reuse and, \textbf{in several cases}, are caused by online...''. A
	definition should be clear - saying in several cases - means that in several
	other cases (1- your several cases) it is not harmful. So, I ask the authors to
	refine and clarify this definition, because as it stands now, this definition is
	confusing (at best). Given that your entire paper is about toxic code snippets,
	you can appreciate why this is important. }

We appreciate the comments. We have refined our definition of toxic code
snippets in the paper by adapting Israel Gat's definition~\citep{toxiccode}
which links toxic code to technical debt, i.e.~``a metaphor reflecting technical
compromises that can yield short-term benefit but may hurt the long-term health
of a software system''~\citep{Li2015}. The new definition of toxic doe snippets
is as follows.

``Toxic code snippets mean code snippets that, after incorporating into
software, the cost of paying back the technical debt exceeds the value it
generates in the long run.'' Stack Overflow code snippets cloned from open
source software or online sources can become toxic when they are (1)~outdated,
(2)~violating their original software license, (3) exhibiting code smells, or
(4) having security vulnerabilities. We chose to study toxic code snippets
caused by outdated code and license-violating code snippets in this study since
these two problems are still underexplored compared to code
smells~\citep{Tufano2015} or vulnerabilities~\citep{Acar2016,Fischer2017}.
Outdated code snippets can be harmful since they are not up-to-date with their
originals and may contain defects. Code snippets from open source projects
usually fall under a specific software license, e.g.,\ GNU General Public License
(GPL). If they are cloned to Stack Overflow answers without the license, and
then flow to other projects with conflicting licenses, legal issues may occur.

\rcomment{ \textbf{RQs need to be better linked:} First, I can see a clear
	motivation for RQs 1-3, however, RQs 4 and 5 seem to be tagged on. I did not see
	a clear transition of why RQs 4 and 5 would follow RQs 1-3. This can be an easy
	fix, but it is an important one since it makes your paper seem less cohesive. }

Regarding RQ4 (Software licensing violation), it needs to follow RQ 1--2 because
the license violation needs evidence of cloning direction. To put it another
way, we could tell that the code snippets on Stack Overflow may have
incompatible licenses to their originals after we confirmed that (1) the Stack
Overflow code snippets were actually clones of Qualitas projects and (2) the
direction of cloning was from Qualitas to Stack Overflow, which both are
addressed in RQ1 and RQ2. Nevertheless, RQ4 does not need to follow RQ3
(Outdated online code clones) because they are independent of each other.

Regarding RQ5 (the Stack Overflow answerer survey), that is a good point. Since
RQ5 supports the motivation of our experiment by showing that code
copying from software projects to Stack Overflow occurs and showing
that some of the answerers are not aware of outdated code and software license
incompatibility. We moved them to RQ1 instead and reorganised the paper to
discuss the survey results as an incentive for our experiment of online code
clone detection and toxic code snippet identification. The findings in RQ2--5
will confirm the survey results by showing that the issues actually occur on
Stack Overflow.

\rcomment{ \textit{Issues related to RQs}

	Since we are talking about the RQs, this might be a good chance to discuss some
	of my other concerns with specific RQs.

	RQ3.~I believe that the answer to this question is expected and perhaps the way
	that the authors answered this question is incorrect. If I use some code from
	SO that was posted in 2014, then I expect that it is outdated. Also, SO does
	not allow people to update their posts, so why would I expect the snippets to
	not be outdated, hence, I feel that your question is incorrect. Another issue
	that might impact the answer to this RQ is the fact that your Qualitas dataset
	is older than that of SO, so your setup encourages the finding of outdated
	snippets. What I think would be a better approach is to look if there are other
	answers (maybe not the accepted one) that are newer. For me, if I see an old
	accepted answer or a newer post with an upvoted answer, I will probably go for
	the newer post. This was never considered by the authors. }

Stack Overflow allows its users to edit their own posts,
i.e.,~questions/answers\footnote{https://meta.stackoverflow.com/questions/315915/i-am-not-allowed-to-edit-my-own-question}, or other user's posts\footnote{https://stackoverflow.com/help/privileges/edit}. Thus, if the code snippet in an answer is outdated and has an issue, the answerer or other high-reputation users can update the answer to be up-to-date or to warn other users that the code is already outdated and may contain problems. We found some discussions regarding this concern on Metaexchange as well so the Stack Overflow community is also considering this option\footnote{Discussion about outdated answer: https://meta.stackexchange.com/questions/131495/how-to-handle-outdated-answers and https://meta.stackexchange.com/questions/180864/editing-outdated-source-code-in-an-answer}. We included this discussion in the paper.

Based on your suggestion of looking at other newer or higher-voted answers, we
performed an investigation of all the answers in 100 posts containing outdated
code snippets as accepted answers. We found that there were 34 posts which
contain newer answers. Regarding the number of votes, there were 5 posts which
contained other answers with a higher number of votes than the outdated answer.
Ninety-nine out of the one hundred outdated code snippets were still marked as accepted
answers. This discussion of the findings is added to RQ4.

\rcomment{ In section 2.4, you consider a change to the code in the Q project to
	be evidence that the snippet is outdated. How do you control for changes that
	are simply due to evolution? For example, what if I refactor the code and change
	the variable names of the snippet. }

Good point. Although we grouped the code modifications that make Stack Overflow
snippets outdated into 6 categories as shown in Table 9, we did not investigate
the motivation of the changes such as code refactoring or variable renaming. We
have performed the check as you suggested by checking the commit message of each
change that makes the code outdated. The findings are discussed below and are
included in RQ4.

We looked through the commit history of 100 outdated code snippets and
summarised the reasons for code updates. We carefully read the commit messages
and/or used the bug IDs in the commit messages to obtain more information from
the bug report systems. We summarised the reasons for code changes into 6 groups
as shown in the table below.

\begin{table}[H]
	\centering
	\begin{tabular}{llr}
		\toprule
		Intent & Detail & Amount \\
		\midrule
		Enhancement & Add or update existing features & 64 \\
		Deprecation & Delete dead/deprecated code & 15 \\
		Bug & Fix bugs & 10 \\
		Refactoring & Refactor code for better design & 6 \\
		Coding style & Update to a new style guideline & 3 \\
		Data change & Changes in the data format & 2 \\
		\bottomrule
	\end{tabular}
	\label{tab:intent_outdated}
	\caption{Intents of code changes in 100 outdated code snippets}
\end{table}

The majority (64) of the changes are due to code enhancement, such as adding
new features or improving existing code to support a newer version of the
language or APIs. Fifteen changes are for removal of dead or deprecated code
such as the SQL connection code example in Hibernate. Ten changes are made to
fix bugs in the software (e.g.,~array reference, race condition, wrong field
access, \texttt{NullPointerException}). There are 6 changes for code
refactoring, 3 changes for updating coding styles using an automated tool or
manually, and 2 changes that are caused by the change of data format.

\rcomment{ \textbf{Data used:} why is the SO data from Jan. 2016? This dump is
	now 2 years old! I can see why you used an older version of the Qualitas
	dataset, but I cannot understand why you would want to use an old version of the
	SO data. Please explain. }

We used the Stack Overflow data dump of January 2016 since it was the latest
dump when we started our experiment. We completed the whole
experiment, the result analysis, and the survey two years later. That is why the
dump is 2 years old.

Fortunately, we could use this opportunity to study the evolution of the toxic
code snippets on Stack Overflow. We manually investigated the outdated and
license incompatible code snippets that we discovered to see whether they are
still the same. The check was performed on 6 May 2018, approximately 2 years
after the snapshot was created. Moreover, we also checked if there is any
comment regarding the outdated code or the license of the snippets from other
Stack Overflow users. From the 100 outdated answers, we found that four answers
were deleted and six answers contained a comment mentioning either the code is
intended for a specific version of the software/API or complaining that the code
is not working or having an issue. We included this discussion in RQ4.

\rcomment{ \textbf{Arbitrary use of parameter:} On page 4, you use snippets of
	at least 10 lines. Why 10, why not 15? Given that this is a journal paper, you
	have the space to examine the impact of your parameters (at least in brief) on
	your results. What I would like the authors to do is to actually examine the
	impact of such parameters on your work in a discussion section.}

That is a fair point. The reason we chose the clone size of at least 10 lines is
because it is considered a standard setting for clone detection in large-scale
data sets to avoid trivial clones (e.g., \texttt{equals}, \texttt{hashCode}, or
getter and setter methods) in the analysis or the manual investigation
phase~\citep{Sajnani2016}.

%Choosing this size threshold larger than 10 could possibly remove some
% non-trivial clones from the analysis, i.e.~create more false negatives.
%On the other hand, decreasing the threshold to be lower than 10 lines will
% result in a larger number of clone candidate pairs to look at and a higher
% chance of getting false positives. Moreover, the large number of clone pairs
% will hinders the full manual validation of the clones and force us to rely on
% sampling of clone pairs. We performed a small analysis to confirm this.
%
%\begin{figure}[H]
%	\centering
%	\includegraphics[width=0.5\linewidth]{../boxplot_clone_size}
%	\caption{Size of online code clones.}
%	\label{fig:boxplotclonesize}
%\end{figure}
%
%Regarding increasing the minimum clone size to more than 10 lines (e.g. 15
% lines), we added an analysis of the clone size as shown in Figure
% \ref{fig:boxplotclonesize}. We found that the distribution of true clones; QS,
% SQ, UD, EX, BP, IN; mainly cover the size from about 10 to 20 lines.
% Increasing the threshold to be higher than 10 would remove some of the trivial
% BP and IN clone pairs which are not interesting for this study. However, it
% will also remove some of the true non-trivial clone pairs in QS, SQ, UD, and EX.
%
%Regarding decreasing the minimum clone size to be less than 10 lines, we reran
% the two clone detection tools (Simian and SourcererCC) again using a minimum
% clone size of 6 lines, another well-accepted minimum clone size in clone
% benchmark~\citep{Bellon2007}. Simian reported 67,172 clone candidate pairs and
% SourcererCC reported 7,752 clone candidate pairs with 776 common clone pairs
% using Bellon's ok-match threshold $t=0.5$. With this large amount of clone
% candidate pairs, it is not feasible to manually confirm all of the clones and
% we must rely on sampling. Moreover, it is known that lowering the number of
% minimum clone lines is harmful to precision. This will result in weaker
% findings and weaker conclusion. Thus, we decide to retain the minimum clone
% size of 10 lines to keep the balance between the number of non-trivial clone
% pairs and the feasibility of manual clone investigation.

To measure the impact of this minimum clone size threshold on the clone results,
we performed an investigation of three different threshold values: six, ten, and
fifteen lines. We started our investigation by using a threshold value of six
lines, a well-accepted minimum clone size in clone benchmark~\citep{Bellon2007}.
Simian reported 67,172 clone candidate pairs and SourcererCC reported 7,752
clone candidate pairs. We randomly sampled 382 pairs from the two sets for
a manual check. This sample number was a statistically significant sample with a
95\% confidence level and $\pm 5\%$ confidence interval. The first author
investigated the sampled clone pairs and classified them into three groups:
non-clones, trivial clones (\texttt{equals}, \texttt{hashCode}, or getter and
setter methods), and non-trivial clones. The manual check found 26 non-clone
pairs; 322 trivial clone pairs, and 34 non-trivial clone pairs. Next, we
increased the threshold to ten lines, another well-established minimum clone
size for large-scale data sets~\cite{Sajnani2016}, and retrieved 721 clone pairs
from Simian and 1,678 clone pairs from SourcererCC. We randomly sampled and
manually checked the same amount of 382 pairs and found 27 non-clone pairs, 253
trivial clone pairs, and 102 non-trivial clone pairs. Then, we increased the
threshold further to fifteen lines and retrieved 196 clone pairs from Simian and
1,230 clone pairs from SourcererCC. The manual check of the 382 randomly sampled
pairs revealed zero non-clone pairs, 298 trivial clone pairs, and 83 non-trivial
clone pairs.

The findings from the three threshold values show that selecting the minimum
clone size of ten lines was preferred over six and fifteen lines. First, it
generated a fewer number of clone pairs than using six lines, which made the
manual clone investigation feasible. Second, it preserved the highest number of
non-trivial clone pairs.

We have modified the paper to include this discussion at the end of Section
2.2.1.

\rcomment{ A similar concern exists on page 5, where t=0.7 is chosen. Sure, it
	worked in another context, but did they show that a 0.7 threshold is OK to use
	for SO snippets? I think you can also vary this t value and examine its impact.
}

We selected Bellon's ok-match threshold of 0.7 because it is a standard value in
clone benchmarking \citep{Bellon2007} and also used in other studies of clones,
e.g.,~\cite{Sajnani2016}. Nevertheless, we agree that varying the threshold ($t$)
value is beneficial to the validity of the findings, so we did an analysis by
choosing five $t$ values of 0.1, 0.3, 0.5, 0.7, and 0.9 and studied the clone
candidates. As shown in our paper, by setting $t=0.7$ according to Bellon's
study, we found 97 ok-match pairs reported. On the other hand, setting $t$ to
0.1, 0.3, 0.5, and 0.9 resulted in 111, 110, 110, and 94 ok-matched pairs
respectively. Since the clone pairs of $t=0.1$ were the superset of other sets, 
we manually checked all the 111 reported pairs. We found one false positive pair
and 110 true positive pairs. By raising the $t$ to 0.3 and 0.5, we got rid of
the false positive pair and still retained all the 110 true positive pairs. All
the clone pairs of $t=0.7$ (97) and $t=0.9$ (94) were also true positives due to
being a subset of $t=0.5$. However, since there were fewer true positive pairs
reported (i.e.,~generating false negatives), we ended up leaving some duplicates
of the same clone pairs (that were supposed to be merged) in the final merged
clone set.

With this analysis, we can see that setting the threshold $t$ to 0.1 is too
relaxed and results in having false positive ok-match pairs while setting the
$t$ to 0.7 or 0.9 is too strict and generates some false negative pairs. Thus,
we decided to select the $t$ value at 0.5. This reduces the number of merged
clone pairs to 2,289 pairs. All other numbers which derive from this merged
clone pairs have been updated accordingly.

\rcomment{ \textbf{Question about clone patterns:} Table 4 lists all the clone
	patterns you considered. First, I have a major concern about the two most
	interesting patterns Q->S and S->Q. Especially in the case of S->Q, how can you
	be sure that this code actually came from SO? It could have come from any other
	project or from SO or another forum. Also, for Q->S, how can you be sure these
	snippets came from the Q projects? This is somewhat of a major concern since the
	rest of your paper is built on this. I would like to see a clear explanation of
	how we can be sure about this.

	\vspace{1ex} Now, you can say that we see code that is similar in Q and S or S
	and Q, but you can (almost) never say that this code comes from SO or from Q
	(to SO). This puts your toxic code idea into question and makes me take the
	results of RQs 1 and 2 with a grain of salt.

	\vspace{1ex} In fact, in the answer of RQ1, you explicitly say ``...share
	similar code with SO ...'', but throughout the paper, you present the work as
	if the code came from SO -> Q or vice versa. }

For the QS clone pairs, as explained in Section 2.2.3 (The online cloning
classification patterns) and Section 3.3 (RQ3: Patterns of Online Code Cloning),
we \textbf{manually confirmed with evidence} that the Stack Overflow code
snippets were actually copied from Qualitas projects. This was done by the two
authors reading the Stack Overflow answers containing the QS code snippets and
looked for evidence such as explanation in the text or the code comments and
made their decision. The two authors also compared their results and discussed
until a consensus was made. Similarly, the same process was applied to the SQ
clone pair. %For the SQ pair we discovered, we found
%evidence of cloning from Stack Overflow post ID 698283 to
%{\small\texttt{POIUtils.java}} in \textsf{jstock} project. The user
%who asked the question on Stack Overflow is an author of
%\textsf{jstock}. The question is about determining the right method to call
%among 7 overloading methods of {\small\texttt{setCellValue}} during runtime. We
%could not find evidence of copying or attribution to Stack Overflow in
%\textsf{jstock}. However, considering that the 25 lines of code of
%{\small\texttt{findMethodToInvoke}} method depicted in Figure 7 in Stack
%Overflow is very similar to the code in \textsf{jstock} including comments, it
% is almost certain that
%the two code snippets form a clone pair. In addition, the Stack Overflow answer
% was
%posted on March 30, 2009, while the code in {\small\texttt{POIUtils}} class in
%\textsf{jstock} was committed to GitHub on the next day of March 31, 2009.
In case of a doubt or not enough cloning evidence and the clone pairs did not
fit into EX, BP and IN, we always categorised the clone pairs into UD (unknown
direction) instead.

To emphasise this point, we also modified the answer of the RQ2 to be clearer.
It now says \textit{``To answer RQ2, we found 2,076 manually confirmed clone
	pairs between 443 Stack Overflow code snippets and 59 Qualitas projects.''.}

\rcomment{ And given the types of clones you find (e.g., BP code), this can
	really come from anywhere (and more likely to come from things like tutorials or
	IDEs). Then, you report that 65\% of the closes are BP. AC is also not a clone
	really since even in your Table 4 you said AC is false clones - how can an
	Accidental *clone* be a false clone? If something is false, then it is not a
	clone. }

You are correct regarding the BP clone pairs. They can really come from anywhere
(we did find that some of the BP clone pairs were auto-generated code by
NetBeans). Therefore, we did not draw any conclusion about them in the paper.
Moreover, thank you for pointing out the inappropriate naming for AC. We now
renamed them from AC to NC (i.e.,~Not Clones).

\rcomment{ I feel that the strongest part of the paper is the survey section. I
	really enjoyed the findings and think that the community can learn something
	from the outcomes of the survey (especially section 3.5.5). However, I did find
	the question in section 3.5.3 to be unfair somehow since there is really no
	mechanism in SO to notify anyone who used a code snippet. In fact, no one even
	knows who used a code snippet from SO. }

%Our apologies if the writing in section 3.5.3 confuses you.
The research question (section 3.5.3) is actually about notifying the answerers
that their answers are already outdated, not notifying the users of the outdated
code snippets. This notification to the answerer can be done by posting comments
for that answer on Stack Overflow. For example, the answerer of the Stack
Overflow question no.~10924648 posted a comment on his own answer saying that
the answer is specifically for Spring 3.1. Moreover, another user posted a
comment on the answer to the post no.~801935 that the code snippet is not of
high quality and suggested an alternative in another post. We extended this
explanation in Section 3.1.1 (RQ1).

\rcomment{ Perhaps minor, but this issue really did not sit well with me, in the
	introduction, you say in RQ5 that you survey 607 participants, but later it
	turns out that only 201 responded. To me, this is somewhat misleading and does
	not sit well. I strongly encourage the authors to refactor such `marketing'
	statements. }

You are right. We fixed the text in the RQ to say that we surveyed 201
participants.

\section{Reviewer 3}

\rcomment{ There is also no discussion if partial updates to the outdate code
	clones where posted in the same Q\&A threat (e.g., addressed by somebody else
	who just mentioned the necessary modifications or provided a link to the updated
	code fragment).  It would also have been interesting to know how old the Q\&As
	(clones) were, since the Qualitas code base is 4 years older then the
	StackOverflow snapshot. Not sure why you selected 2013 and not 2014 or 2012.
	Average and mean time/date of the clone postings on stackoverflow would be
	interesting to see the distribution of the clone age (simple boxplot would do).
}

We are grateful for the insightful comment. We selected the 2013 snapshot of
Qualitas data since it is the latest and the most complete snapshot (i.e.,~having
the highest number of Java projects) provided by the data set author.

Regarding the age of cloned code on Stack Overflow, we did an analysis of clone
ages as you suggested. We only chose the QS clone pairs since we could manually
confirm the direction of cloning. Moreover, we knew the date that the original
version of the QS clones appeared in Qualitas projects (we used the version
release date), which was not possible for UD, EX, BP, and IN clone pairs. We
calculated the clone ages by counting the number of months between the date of
each Qualitas project and the date the answer was posted on Stack Overflow.

\begin{figure}[H] \centering
	\includegraphics[width=0.7\linewidth]{../boxplot_clone_age} \caption{Age of QS
		cloned snippets} \label{fig:boxplotcloneage} \end{figure}

We drew a boxplot as you suggested (shown above) and included the discussion of
the statistics in RQ4. We found that the minimum, maximum, median, and average
clone ages were 0, 64, 26.5 and 26 months respectively. We observed that, on
average, it took around 2 years for the clones from Qualitas projects to appear
on Stack Overflow answers. Some of the clones appeared on Stack Overflow almost
at the same time as the original, while the oldest clones took around 5 years.

\rcomment{ I do believe the survey results from the StackOverflow contributors
	actually provides more meaningful insights than the clone detection itself,
	since the responses by the contributors actually help explaining some of the
	results. The authors should provide further analysis and interpretation of the
	survey results. }

We moved them to RQ1 to better link with the motivation. We also expanded the
findings to include another survey, the Stack Overflow visitor survey,
to understand the awareness of the developers on both sides: the answerers and
the visitors (i.e.,~information seekers). We asked the visitors their experience
regarding outdated and license-incompatible code caused by copying and reusing
code snippets from Stack Overflow. The findings are incorporated into RQ1.

\rcomment{ Overall discussion is very weak. While the authors went through a
	lengthy and detailed study of code clones published on StackOverflow and the
	survey of StackOverflow contributors, the paper lacks really a take home
	message. At the end there is no real evidence that either code clones are really
	toxic or bad (are they actually causing problems? due to the use of clones. Just
	because a code fragment is outdated does not mean it is necessarily bad or
	causes any problems. What is the actual take home message from your study?
	Should we stop reusing code from StackOverflow? I do believe there is a need to
	dig deeper in the observed data and draw some real conclusions on how the
	observed results can benefit both the StackOverflow community and the source
	code community. }

That is a good point. We relied on the two new analyses and added a new
discussion about the toxicity of outdated code in the Overall Discussion.

First, from the answer to Reviewer 2, we analysed the intent behind the changes
that made the code outdated and found that only 10 out of 100 cases were bug
fixes. Besides that, the code snippets were outdated because of evolution in the
software (adding or enhancing features, refactoring, changing coding style, and
data change). Thus, outdated code can sometimes become toxic, but not always.

Second, as previously discussed in answer to Reviewer 1, we have performed
an additional experiment to search for occurrences of the 100 outdated code
snippets in 130,703 open source GitHub projects and found 102 clones of 30
outdated code snippets. 47 were an exact copy of the outdated versions, of which 12
were buggy. However, we could not establish evidence that the outdated code
snippets were cloned from Stack Overflow or somewhere else. We have reported the
problem of race condition in two clone snippets found in deeplearning4j and
Apache Hive to the developers.

To summarise, we empirically show from the surveys and the clone detection
experiment that online code clones occur on Stack Overflow and the clones may
become toxic due to outdated code and software license incompatibility. For
outdated code, we found only a small number of clones. Although other studies
show that Stack Overflow code snippets may contain security
vulnerabilities~\citep{Acar2016,Fischer2017} or API misuse~\citep{Zhang2018}, we
found in this study that the damage caused by reusing outdated code on Stack
Overflow is not high. On the other hand, the missing licensing statements of
online code clones on Stack Overflow can be more harmful. As shown in our study
and also in the study by \cite{An2017}, some online clones on Stack Overflow are
originally licensed under more restrictive license than Stack Overflow's CC
BY-SA 3.0. If these missing-license online clones are reused in software with
an incompatible license, the software owner may face legal issues. Software
auditing services such as Black Duck Software or nexB, which can effectively
check for license compliance of code copied from open source projects, do not
check for the original license of the code snippets on Stack Overflow. Although
the Stack Overflow answerers who participated in our survey believe that most of
the code snippets on Stack Overflow are too small to claim for copyright and
they fall under fair-use, there is still a risk due to different legal systems
in each country. For example, Germany's legal system does not have a concept of
fair use or the number of minimum lines of code to be considered trivial,
i.e.,~de minimis, is also different from case to case or from country to country.

We also updated the Actionable Items section to reflect Reviewer 1's comment.

\vspace{1cm}

\section{Conclusion}
We would like to thank the reviewers again for their comments.
We have tried our best to address every concern that has been pointed out by the reviewers. This result in inclusion of a Stack Overflow visitors' survey, an analysis of toxic code snippets in GitHub projects, analysis of different tools' and technique's parameters, a more detailed analysis of online code clones, the updated definition of toxic code snippets, the improved overall discussion section, and other minor improvements throughout the paper.
We believe this new submission has been considerably improved from the previous version due to the insightful reviews and discussions and we are looking forward to the new reviews.

% Uncomment in case references are needed
%\bibliographystyle{plain}
\bibliographystyle{spbasic}
\bibliography{references}

\end{document}